---
title: " Analisi Schede di Dimissioni Ospedaliere "
author: "| Marco Gaudio |"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    highlight: tango
    keep_md: no
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  github_document:
    toc: yes
---

```{r, include=FALSE}
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE, 
                      include = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = FALSE)
library(tidymodels)
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(DT)
library(lemon)
```


The aim of this document is to analyze the data related to the italian *Schede di Dimissioni Ospedaliere* (SDO), in the period 2016 - 2019. 
The data are provided by the [**Italian Ministry of Health**](http://www.salute.gov.it/portale/temi/p2_6.jsp?lingua=italiano&id=1237&area=ricoveriOspedalieri&menu=vuotolink).

The document is composed by 4 sections: 

* **section 1** provides an analysis for each of the 5 activities (*Acuti Regime ordinario*; *Acuti Regime diurno*, *Riabilitazione Regime ordinario*; *Riabilitazione Regime diurno*; *Lungodegenza*) and considering the differences in the period studied. 

* **section 2** analyses the hospitalization rates, first providing an overview for Italy as a whole, and then analysing the rates for each of the 5 activities in in the period 2016-2019, considering different regions, gender and age categories.

* **section 3** gives an overview of the Italian and English population in 2019, then an analysis of England's total number of surgeries (that can be compared to the total number of the italian **"dimissioni"**) and surgical incidences (the italian **"tasso di ospedalizzazione"**) is made. Lastly, it shows a comparison between the overall hospitalization rates in Italy and England, thanks to the data provided by the [**Health and Social Care Information Centre**](https://digital.nhs.uk/), now called NHS Digital. 

* **section 4** proposes an analysis of the data about hospitalization discharges collected from two different sources: [**Istat**](http://dati.istat.it/) and Italian Ministry of Health.


```{r}
# construct a summary table
summary_ordinario <- matching_tables %>% 
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Acuti - Regime ordinario")

summary_diurno <- matching_tables_2.2.11 %>% 
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Acuti - Regime diurno")

summary_riab_ord <- SDO_ordinario_riab %>%
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Riabilitazione - Regime ordinario")

summary_riab_diurno <- SDO_diurno_riab %>%
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Riabilitazione - Regime diurno")

summary_lungodegenza <- SDO_lungodegenza %>%
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Lungodegenza")

summary_tab <- dplyr::bind_rows(summary_ordinario,
                                summary_diurno,
                                summary_riab_ord,
                                summary_riab_diurno,
                                summary_lungodegenza)

```


### An overview of hospitalization discharges in Italy {-}
The following graph shows the total number of hospital discharges divided by activity for the 4 years of analysis.

```{r, echo=FALSE, include=TRUE}

  p <- ggplot(summary_tab, aes(x = ANNO, y = DIMISSIONI, fill = `ATTIVITÀ`)) + 
  geom_bar(stat = "identity") +
  labs(title = "DIMISSIONI TOTALI PER ATTIVITÀ",
  subtitle = "Italia",
  caption = "Data Source: Ministero della Salute")
p
```

Table 1.1.1 shows the total number of discharges for all the activities and its percentage variation through the period of analysis. As shown here, the total number of hospital discharges is decreased from 2016 to 2019. In particular, from 2018 to 2019 the number has dropped from 8.356.874 to 8.212.031, with a decrease of 1.733%.

```{r, echo=FALSE, include=TRUE}

summary_tab %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) 
                                       *100, digits = 3)) %>%
DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.1.1')

```



# ATTIVITÀ

## Acuti - Regime ordinario.
the following section is aimed to analyze the discharge in *Acuti - Regime ordinario*, from 2016 to 2019 in Italy.

### DIMISSIONI

Table 1.1.2 provides a summary of the hospedalization discharges in *"Acuti - Regime ordinario"* and the percentage variation year by year.

```{r, echo=FALSE, include=TRUE}
summary_ordinario <- matching_tables %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_ordinario %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.1.2')



```

This bar chart shows the hospital discharges by [MDC](#glossario) for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_ordinario %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER MDC",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute") + 
  xlab("CLASSE MDC")

  
```

As shown from the bar chart, over the 4 years of analysis, the MDC class with the greatest number of discharges for the activity *Acuti - Regime ordinario* is the **MDC 05 - Malattie e disturbi dell'apparato cardiocircolatorio**, followed by **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo** and **MDC 04 - Malattie e disturbi dell'apparato respiratorio**.  



The following table provides a summary of the discharges for each MDC class and for each year.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime ordinario for the 4 years
overall_ordinario <- matching_tables %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_ordinario %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.1.3')
```

### PERCENTUALE GIORNI DIMISSIONI PER MDC.

The following pie charts show the number of days before discharge in percentage for the 4 years analysed. As shown, the highest percentage of discharges for the activity *Acuti - Regime ordinario* is *DIMISSIONI TRA 4 GIORNI E VALORE SOGLIA*, with 55.6%. This means that 55.6% of the hospitalization discharges in this activity lasts between 4 days and the treshold value ([**Valore Soglia**](#glossario)).


```{r, echo=FALSE, include=TRUE}

percentage_tab <- function(year){
  perc <- SDO_ordinario %>%
  filter(ANNO == year) %>%
  mutate(percentage_dim1gg = `% DIMISSIONI DI 1 GIORNO`/100,
         percentage_dim2gg = `% DIMISSIONI DI 2-3 GIORNI`/100,
         percentage_dim4gg = `% DIMISSIONI TRA 4 GIORNI E VALORE SOGLIA`/100,
         percentage_dimooltre = `% DIMISSIONI OLTRE VALORE SOGLIA`/100) 
# compute the percentage for the whole activity

dimissioni <- as.matrix(perc$DIMISSIONI, ncol = 1)  
percent_oneday <-  as.matrix(perc$percentage_dim1gg, nrow = 1) 
percent_twoday <- as.matrix(perc$percentage_dim2gg, nrow = 1)
percent_fourday <- as.matrix(perc$percentage_dim4gg, nrow = 1)
percentbeyond <- as.matrix(perc$percentage_dimooltre, nrow = 1)

dimissioni_1 <- round((t(percent_oneday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_2 <- round((t(percent_twoday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_4 <- round((t(percent_fourday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_oltre <- round((t(percentbeyond) %*% dimissioni )/ sum(perc$DIMISSIONI), 
                          digits = 3)
return(c(dimissioni_1,dimissioni_2, dimissioni_4, dimissioni_oltre))
}


mycols <- c("% DIMISSIONI DI 1 GIORNO",                 
            "% DIMISSIONI DI 2-3 GIORNI",              
            "% DIMISSIONI TRA 4 GIORNI E VALORE SOGLIA",
            "% DIMISSIONI OLTRE VALORE SOGLIA")

# function to create a table with data to plot in a pie chart
prop_table <- function(year){
  discharges_prop <- percentage_tab(year)
tabella_plot <- tidyr::tibble(class = mycols,
                       proportion = discharges_prop) %>%
  dplyr::arrange(-proportion) %>%
  dplyr::mutate(Per_cumsum = cumsum(proportion)) %>%
  dplyr::mutate(ANNO = year)

}
Years <- c("2016", "2017", "2018", "2019")

summary_list <- purrr::map(Years, prop_table)

summary_prop <- dplyr::bind_rows(summary_list) 

ggplot(data = summary_prop, aes(x=1, y=proportion, fill=class)) +
  geom_col() +
  geom_text(aes(label = proportion*100), position = position_stack(vjust = 0.5))+
  coord_polar(theta = "y") + 
  theme_void() +
  labs(title = "% DIMISSIONI PER GIORNO",
       subtitle = "Acuti - Regime ordinario",
       fill = "",
       caption = "Data Source: Ministero della Salute") +
  facet_wrap(~ANNO)



```






The bar chart below shows the distribution of discharges in *Acuti - Regime ordinario* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime ordinario") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

# re-level factor
sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ",
       subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute"
       ) + 
  ylab("DIMISSIONI") + 
  xlab("CATEGORIA ETÀ")

  
  




```

The following table provides a summary of the data presented about hospitalization discharges
by age category.


```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 %>%
  dplyr::select(-TAVOLA) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.1.4')

```



### GIORNATE DI DEGENZA {.tabset}

The following graph shows the [*"Giorni di Degenza"*](#glossario) for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_ordinario %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(GIORNI_DEGENZA = sum(`GIORNI DI DEGENZA`)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = GIORNI_DEGENZA,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "GIORNI DI DEGENZA",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute") + 
  ylab("GIORNI DI DEGENZA") + 
  xlab("CLASSE MDC")


```


Table 1.1.5 provides a summary for *Giorni di Degenza* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_ordinario_degenza <- matching_tables %>%
  dplyr::select(CLASSE_MDC, `GIORNATE DI DEGENZA`, ANNO)

overall_ordinario_degenza %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 1.1.5"
                )

```

The following graph shows the Mean value for *Giorni di Degenza*, for each MDC and for each year.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```

#### 2017 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```

### {-}

Table 1.1.6 provides a summary of Max, Min and Mean values for *"Giorni di Degenza"* for each MDC.
```{r, echo=FALSE, include=TRUE}

SDO_ordinario %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3)) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.1.6'
    
  )

```



## Acuti - Regime diurno.

### DIMISSIONI

Table 1.2.1 provides a summary of the hospedalization discharges in *"Acuti - Regime diurno"* and the percentage variation year by year. As shown, there has been a decrease in the hospitalization discharges in this activity, in particular from 2016 to 2017, with a reduction of 6.71%.


```{r, echo=FALSE, include=TRUE}
summary_diurno <- matching_tables_2.2.11 %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_diurno %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.2.1')



```


This bar chart shows the hospital discharges by MDC for each year of the analysis.

```{r, echo=FALSE, include=TRUE}

SDO_diurno %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER MDC",
              subtitle = "Acuti - Regime diurno",
       caption = "Data Source: Ministero della Salute") + 
  xlab("CLASSE MDC")

  
```


In the case of *Acuti - Regime diurno*, over the 4 years of analysis, the MDC class with the greatest number of discharges is the **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**, followed by **MDC 17 - Malattie e disturbi mieloproliferativi e neoplasie scarsamente differenziate** and **MDC 13 -Malattie e disturbi dell'apparato riproduttivo femminile**.  


The following table provides a summary of the discharges for each MDC class and for each year.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime diurno for the 4 years
overall_diurno <- matching_tables_2.2.11 %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_diurno %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.2.2')
```

The bar chart below shows the distribution of discharges in *Acuti - Regime diurno* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime diurno") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ",
       subtitle = "Acuti - Regime diurno",
       caption = "Data Source: Ministero della Salute") + 
  ylab("DIMISSIONI") + 
  xlab("CATEGORIA ETÀ")

  
  




```

The following table provides a summary of the data presented about hospitalization discharges
by age category.


```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 %>%
  dplyr::select(-TAVOLA) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.2.3')

```



### ACCESSI {.tabset}

The following graph shows the [*Accessi*](#glossario) for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_diurno %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(Accessi = sum(Accessi)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = Accessi,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "ACCESSI",
              subtitle = "Acuti - Regime diurno",
       caption = "Data Source: Ministero della Salute") +
  ylab("ACCESSI") + 
  xlab("CLASSE MDC")


```

These bar charts shows the mean for the variable *Accessi* by each MDC and for the 4 years of analysis.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  

```


#### 2018 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  

```



#### 2017 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       x = "CLASSE MDC",
       caption = "Data Source: Ministero della Salute",
       y = "ACCESSI") 
  

```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       x = "CLASSE MDC",
       caption = "Data Source: Ministero della Salute",
       y = "ACCESSI") 
  


```




### {-}

Table 1.2.4 provides a summary of Max, Min and Mean values for *"Accessi"* for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_diurno %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3)) %>% 
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.2.4'
    
  )
```


## Riabilitazione - Regime ordinario.

### DIMISSIONI

Table 1.3.1 provides a summary of the hospedalization discharges in *"Riabilitazione - Regime ordinario"* and the percentage variation from year to year. There is a reduction in the number of discharges from 2016 to 2018, then a slight increase in 2019. None of these is very significant.

```{r, echo=FALSE, include=TRUE}
summary_ordinario_riab <- SDO_ordinario_riab %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100,
                                       digits = 3))

summary_ordinario_riab %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.3.1')



```

This bar chart shows the hospital discharges by MDC for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute") + 
  xlab("CLASSE MDC")

  
```

There are just a few MDC classes with a considerable number of discharges in *Riabilitazione - Regime ordinario*. These are: **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**; **MDC 01 - Malattie e disturbi del sistema nervoso**; **MDC 05 - Malattie e disturbi dell'apparato cardiocircolatorio**.


Table 1.3.2 provides a summary of the discharges for each MDC class and for each year.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime diurno for the 4 years
overall_riab_ord <- SDO_ordinario_riab %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_riab_ord %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.3.2')
```


The bar chart below shows the distribution of discharges in *Riabilitazione - Regime ordinario* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione - Regime ordinario") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ",
       subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute") +
  ylab("DIMISSIONI") + 
  xlab("CATEGORIA ETÀ")

  

```


The following table provides a summary of the data presented about hospitalization discharges
by age category.


```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 %>%
  dplyr::select(-TAVOLA) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.3.3')

```

### GIORNATE DI DEGENZA {.tabset}

The following graph shows the *Giorni di Degenza* for each MDC and over the years of analysis.

```{r, echo=FALSE, include=TRUE}
SDO_ordinario_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(GIORNI_DEGENZA = sum(`GIORNATE DI DEGENZA`)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = GIORNI_DEGENZA,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "GIORNI DI DEGENZA",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute") + 
  ylab("GIORNI DI DEGENZA") + 
    xlab("CLASSE MDC")


```


Table 1.3.4 provides a summary for *Giorni di Degenza* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_ordinario_degenza_riab <- SDO_ordinario_riab %>%
  dplyr::select(CLASSE_MDC, `GIORNATE DI DEGENZA`, ANNO)

overall_ordinario_degenza_riab %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 1.3.4"
                )

```



The following graph shows the Mean value for *Giorni di Degenza* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2017 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

### {-}

Table 1.3.5 provides a summary for the mean value for *"Giorni di Degenza"*, for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::select(CLASSE_MDC, `DEGENZA MEDIA (giorni)`, ANNO) %>%
  mutate_at(vars(`DEGENZA MEDIA (giorni)`), funs(round), digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.3.5'
    
  )
  


```


## Riabilitazione - Regime diurno.

### DIMISSIONI

Table 1.4.1 provides a summary of the hospedalization discharges in *"Riabilitazione - Regime diurno"* and the percentage variation from year to year. As shown, there has been an important decrease in hospitalization discharges in between 2017 and 2018 (-6.745%).

```{r, echo=FALSE, include=TRUE}
summary_diurno_riab <- SDO_diurno_riab %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100,
                                       digits = 3))

summary_diurno_riab %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.4.1')


```



This bar chart shows the hospital discharges by MDC for each year of the analysis.

```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute") + 
  xlab("CLASSE MDC")

  
```

In the case of activity *Riabilitazione - Regime diurno*, there are a few MDC with a considerable level of discharges: **MDC 01 - Malattie e disturbi del sistema nervoso**;
**MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**; **MDC 23 - Fattori che influenzano lo stato di salute ed il ricorso ai servizi sanitari**.

The following table provides a summary of the discharges for by MDC class and for each year of the analysis.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime diurno for the 4 years
overall_diurno_riab <- SDO_diurno_riab %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_diurno_riab %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.4.2')
```


The bar chart below shows the distribution of discharges in *Riabilitazione - Regime diurno* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione - Regime diurno") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ",
       subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute") + 
  xlab("CATEGORIA ETÀ") + 
  ylab("DIMISSIONI")

  

```

The following table provides a summary of the data presented about hospitalization discharges
by age category.


```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 %>%
  dplyr::select(-TAVOLA) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.4.3')

```


### ACCESSI {.tabset}

The following graph shows the *Accessi* for each MDC and over the years of analysis.

```{r, echo=FALSE, include=TRUE}
SDO_diurno_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(Accessi = sum(ACCESSI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = Accessi,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "ACCESSI",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute") + 
  ylab("ACCESSI") + 
  xlab("CLASSE MDC")


```


Table 1.4.4 provides a summary for *Accessi* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_diurno_accessi_riab <- SDO_diurno_riab %>%
  dplyr::select(CLASSE_MDC, ACCESSI, ANNO)

overall_diurno_accessi_riab %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 1.4.4"
                )

```



The following graph shows the Mean value for *Accessi* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```


#### 2017 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```



#### 2016 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```


### {-}


Table 1.4.5 provides a summary for the mean value for *Accessi* for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::select(CLASSE_MDC, `NUMERO MEDIO DI ACCESSI`, ANNO) %>%
  mutate_at(vars(`NUMERO MEDIO DI ACCESSI`), funs(round), digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.4.5'
    
  )
  


```


## Lungodegenza.

### DIMISSIONI

Table 1.5.1 provides a summary of the hospedalization discharges in [*Lungodegenza*](#glossario) and the percentage variation year by year.

```{r, echo=FALSE, include=TRUE}
summary_lungodeg <- SDO_lungodegenza %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_ordinario_riab %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.5.1')



```



This bar chart shows the hospital discharges by MDC for each year of the analysis.



```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER MDC",
              subtitle = "Lungodegenza", 
       caption = "Data Source: Ministero della Salute") + 
  xlab("CLASSE MDC")

  
```

As shown from the bar chart, over the 4 years of analysis, the MDC classes with the greatest number of discharges for the activity *Lungodegenza* are: **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**; **MDC 01 - Malattie e disturbi del sistema nervoso**; **MDC 04 - Malattie e disturbi dell'apparato respiratorio**.

The following table provides a summary of the discharges for each MDC class and for each year.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime diurno for the 4 years
overall_lungod <- SDO_lungodegenza %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_lungod %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.5.2')
```

The bar chart below shows the distribution of discharges in *Lungodegenza* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Lungodegenza") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ",
       subtitle = "Lungodegenza",
       caption = "Data Source: Ministero della Salute") +
  xlab("CATEGORIA ETÀ") + 
  ylab("DIMISSIONI")

  

```


The following table provides a summary of the data presented about hospitalization discharges
by age category.


```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 %>%
  dplyr::select(-TAVOLA) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.5.3')

```
### GIORNATE DI DEGENZA {.tabset}

The following graph shows the *Giorni di Degenza* for each MDC and over the years of analysis.

```{r, echo=FALSE, include=TRUE}
SDO_lungodegenza %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(GIORNI_DEGENZA = sum(`GIORNATE DI DEGENZA`)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = GIORNI_DEGENZA,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "GIORNI DI DEGENZA",
              subtitle = "Lungodegenza",
       caption = "Data Source: Ministero della Salute") + 
  ylab("GIORNI DI DEGENZA") + 
  xlab("CLASSE MDC")


```


Table 1.5.4 provides a summary for *Giorni di Degenza* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_lungod <- SDO_lungodegenza %>%
  dplyr::select(CLASSE_MDC, `GIORNATE DI DEGENZA`, ANNO)

overall_lungod %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 1.5.4"
                )

```



The following graph shows the Mean value for *Giorni di Degenza* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC", 
       caption = "Data Source: Ministero della Salute",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       caption = "Data Source: Ministero della Salute",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2017 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       caption = "Data Source: Ministero della Salute",
       y = "GIORNI DI DEGENZA") 
  




```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_bar(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       caption = "Data Source: Ministero della Salute",
       y = "GIORNI DI DEGENZA") 
  




```

### {-}

Table 1.5.4 provides a summary for the mean value for *"Giorni di Degenza"* for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::select(CLASSE_MDC, `DEGENZA MEDIA (giorni)`, ANNO) %>%
  mutate_at(vars(`DEGENZA MEDIA (giorni)`), funs(round), digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.5.5'
    
  )
  


```



# TASSI DI OSPEDALIZZAZIONE 

## Overview in Italia

### TASSO DI OSPEDALIZZAZIONE IN ITALIA PER ATTIVITÀ
The following bar chart shows the evolution of the [*Tasso di ospedalizzazione*](#glossario) in Italy in the period of analysis.
As shown by the graph, there has been a decrease in the hospitalization rates per 1000 inhabitants, in particular for the activity *"Acuti - Regime ordinario"* (from 101.624 in 2016 to 97.611 in 2019).

```{r, echo=FALSE, include=TRUE}
HR_summary_act <- HR_age %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITALIA") %>%
  dplyr::group_by(ANNO, `ATTIVITÀ`) %>%
  dplyr::mutate(`ATTIVITÀ` = stringr::str_remove_all(string = `ATTIVITÀ`,
                                                     pattern = "Regime")) %>%
  dplyr::select(`Tasso complessivo`, ANNO, `ATTIVITÀ`) 

  ggplot(data = HR_summary_act, 
         mapping = aes(x = `ATTIVITÀ`, 
                       y = `Tasso complessivo` ,
                       fill = ANNO)) +
  geom_bar(position = "dodge",
           width = .5,
           stat = "identity", 
           color = "black") + 
  coord_flip() +
  theme(axis.text.x  = element_text(angle = 90, 
                                     hjust = 1, 
                                    vjust = 0.5)) + 
   labs(title = "TASSO DI OSPEDALIZZAZIONE (per 1000 abitanti)",
              subtitle = "ITALIA",
        caption = "Data Source: Ministero della Salute") + 
    xlab("TASSO")
  
```

Table 2.1.1 provides an overview on the hospitalization rates (per 1000 inhabitants) in Italy during the period 2016-2019.

```{r, echo=FALSE, include=TRUE}

HR_age %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITALIA") %>%
  dplyr::rename(`Tasso complessivo Italia` = "Tasso complessivo") %>%
  dplyr::select(`ATTIVITÀ`, `Tasso complessivo Italia`, ANNO) %>%
  dplyr::mutate_at(vars(`Tasso complessivo Italia`), funs(round), digits = 3) %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 2.1.1"
  )

```

### TASSI DI OSPEDALIZZAZIONE PER GENERE

The graph below shows the hospitalization rates in Italy per gender and for the different activities. In this case, there has been a small reduction for both gender and for all the activities.

```{r, echo=FALSE, include=TRUE}
gender_OR <- HR_gender %>%
  dplyr::group_by(ANNO, `ATTIVITÀ`) 


                                          
# reshape the dataframe in order to plot.

male <- gender_OR %>%
  dplyr::select(1, 2, 4, 5) %>%
  dplyr::rename(TASSO = "MASCHI") %>%
  dplyr::mutate(GENERE = "MASCHI") 

female <- gender_OR %>%
  dplyr::select(1,3, 4, 5) %>%
  dplyr::rename(TASSO = "FEMMINE") %>%
  dplyr::mutate(GENERE = "FEMMINE") 


Gend_HR_total <- dplyr::bind_rows(male, female) %>%
  dplyr::mutate_at(vars(TASSO), as.numeric)

Italy_HR <- Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITALIA") %>%
  dplyr::rename(tasso_it = "TASSO") %>%
  dplyr::mutate(`ATTIVITÀ` = stringr::str_remove_all(string = `ATTIVITÀ`,
                                                     pattern = "Regime")) 
  
  ggplot( data = Italy_HR, 
          aes(x =`ATTIVITÀ`, y = tasso_it, fill = GENERE)) +
  geom_bar(
    stat = "identity",
    position = "dodge"
    ) +
    coord_flip() +
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (per genere, per 1000 abitanti)",
              subtitle = "Italia",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") + 
  facet_wrap(~ANNO)
  


```



Table 2.1.2 provides a summary for the data about hospitalization rates by gender in Italy.


```{r, echo=FALSE, include=TRUE}

Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITALIA") %>%
  dplyr::rename(tasso_it = "TASSO") %>%
  dplyr::select(-1) %>%
  dplyr::mutate_at(vars(tasso_it), round, digits = 3) %>%
  dplyr::rename(`Tasso osp. per 100 abitanti` = "tasso_it") %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 2.1.2"
  )


```



### TASSO DI OSPEDALIZZAZIONE PER CATEGORIA DI ETÀ ED ATTIVITÀ

The following graph represents the evolution of the hospitalization rates by age category in Italy and for each activity.

```{r, echo=FALSE, include=TRUE}

# write a function to reshape the dataset
convert_HR_age <- function(activity){
  
  HR_summary_it <- HR_age_cat %>%
    dplyr::select(`Categoria età`, activity, ANNO) %>%
    dplyr::rename(TASSO = activity) %>%
    dplyr::mutate(`ATTIVITÀ` = activity) %>%
    dplyr::slice(-n())
  HR_summary_it$`Categoria età` <- factor(HR_summary_it$`Categoria età`,
                                      c("Meno di 1 anno",
                                        "Da 1 a 4 anni",
                                        "Da 5 a 14 anni",
                                        "Da 15 a 24 anni",
                                        "Da 25 a 44 anni", 
                                        "Da 45 a 64 anni", 
                                        "Da 65 a 74 anni", 
                                        "75 anni e oltre"))
  return(HR_summary_it)
}

activity_list <- list("Acuti - Regime ordinario", "Acuti - Regime diurno",
                      "Riabilitazione - Regime ordinario", 
                      "Riabilitazione - Regime diurno",
                      "Lungodegenza")

list_HR_age <- purrr::pmap(list(activity_list), convert_HR_age)

HR_age_it <- dplyr::bind_rows(list_HR_age) %>%
  dplyr::filter(!is.na(`Categoria età`)) %>%
  dplyr::group_by(ANNO, `ATTIVITÀ`) 

  ggplot(data = HR_age_it, aes(x = `Categoria età`,
                                   y = `TASSO`,
                                   fill = `ATTIVITÀ`,
                                   )) +
  geom_bar(stat = "identity",
           position = "dodge") +
    theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (PER ETÀ, per 1000 abitanti)",
              subtitle = "Italia",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") +
    xlab("CATEGORIA ETÀ")+
  facet_wrap(~ANNO)
  
     

```



Table 2.1.3 provides the summary for the hospitalization rates in Italy by age category. 


```{r, echo=FALSE, include=TRUE}

HR_age_it <- dplyr::bind_rows(list_HR_age) %>%
  dplyr::filter(!is.na(`Categoria età`)) %>%
  dplyr::mutate_at(vars(TASSO), round, digits = 3)

HR_age_it %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 2.1.3"
  )

```


## Acuti - Regime ordinario

Table 2.2.1 shows the hospitalization rates for the activity *Acuti - Regime ordinario* for the 4 years of analysis and the corresponding percentage variation. A 1.998% reduction in the hospitalization rates occurred between 2018 and 2019.

```{r, echo=FALSE, include=TRUE}
HR_summary_act %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti -  ordinario") %>%
  dplyr::mutate(`ATTIVITÀ` = "Acuti - Regime ordinario") %>%  
  dplyr::rename(TASSO = "Tasso complessivo") %>%
  dplyr::mutate_at(vars(TASSO), round, digits = 3) %>%
  ungroup() %>%
  dplyr::mutate("VARIAZIONE %" = round((TASSO / lag(TASSO) -1) * 100, 
                                       digits   = 3)) %>%
  DT::datatable(
    options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 2.2.1')
  

```


### TASSO PER CATEGORIE DI ETÀ

As shown by the graph, through the years of analysis the hospitalization rates by age category have a little drop for the activity *Acuti - Regime ordinario*.

```{r, echo=FALSE, include=TRUE}

HR_summary <- HR_age_cat %>%
  dplyr::filter(`Categoria età` != "TOTALE") %>%
  dplyr::select(`Categoria età`, `Acuti - Regime ordinario`, ANNO) %>%
  dplyr::group_by(ANNO) 
  
HR_summary$`Categoria età` <- factor(HR_summary$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 14 anni",
                                            "Da 15 a 24 anni",
                                            "Da 25 a 44 anni", 
                                            "Da 45 a 64 anni", 
                                            "Da 65 a 74 anni", 
                                            "75 anni e oltre"))

  ggplot(data = HR_summary,
         aes(x = `Categoria età`, y = `Acuti - Regime ordinario`, fill = ANNO)) +
  geom_bar( stat = "identity",
            position = "dodge") +
     theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (per età, per 1000 abitanti)",
              subtitle = "Acuti - Regime ordinario",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") + 
    xlab("CATEGORIA ETÀ")


```



Table 2.2.2 represents the data about hospitalization rates for age category in *Acuti - Regime ordinario* for each year.


```{r, echo=FALSE, include=TRUE}

HR_summary %>%
  dplyr::rename(`Tasso di ospedalizz. (per 1000 ab.)` = "Acuti - Regime ordinario") %>%
  dplyr::mutate_at(vars(`Tasso di ospedalizz. (per 1000 ab.)`), round, digits = 3) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 2.2.2"
                )

  

```


### TASSO PER REGIONE E SESSO

The following graphs show the evolution of the hospitalization rate in Italy by region and gender in the period 2016-2019. The horizontal blue line displays the hospitalization rates in all Italy for the male gender, the red one for the female gender.

```{r, echo=FALSE, include=TRUE}

gender_OR <- HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime ordinario") %>%
  dplyr::group_by(ANNO) 

# change the regions names in shorter ones.
gender_OR$`REGIONE DI RESIDENZA` <- rep(c("Pie", "VdA", "Lom", "Blz", "Trn", "Ven",
                                      "FVG", "Lig", "EmR", "Tos", "Umb", "Mar",
                                      "Laz", "Abr", "Mol", "Cam", "Pug", "Bas",
                                      "Cal", "Sic", "Sar", "ITA"),4
                                      )
gender_OR$`REGIONE DI RESIDENZA` <- factor(gender_OR$`REGIONE DI RESIDENZA`,
                                          c( "Pie", "VdA", "Lom", "Blz",
                                             "Trn", "Ven", "FVG", "Lig", 
                                             "EmR", "Tos", "Umb", "Mar",      
                                             "Laz", "Abr", "Mol", "Cam", 
                                             "Pug", "Bas","Cal", "Sic",
                                             "Sar", "ITA")
                                          )
# reshape the dataframe in order to plot.

male <- gender_OR %>%
  dplyr::select(1, 2, 5) %>%
  dplyr::rename(TASSO = "MASCHI") %>%
  dplyr::mutate(GENERE = "MASCHI") 

female <- gender_OR %>%
  dplyr::select(1,3,5) %>%
  dplyr::rename(TASSO = "FEMMINE") %>%
  dplyr::mutate(GENERE = "FEMMINE") 


Gend_HR_total <- dplyr::bind_rows(male, female) %>%
  dplyr::mutate_at(vars(TASSO), as.numeric)

Italy_HR <- Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITA") %>%
  dplyr::rename(tasso_it = "TASSO")

Gend_HR_without_ita <- Gend_HR_total %>%
  filter(`REGIONE DI RESIDENZA` != "ITA")


p <- ggplot(data = Gend_HR_without_ita, aes( x = `REGIONE DI RESIDENZA`, 
                                 y = TASSO,
                                 fill = GENERE)) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "MASCHI"), aes(yintercept = tasso_it), 
             color = "blue") +
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "FEMMINE"), aes(yintercept = tasso_it), 
             color = "red") +
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +

  facet_wrap(~ANNO) +
  labs(
    title = "Tasso di ospedalizzazione per regione e sesso (per 1000 abitanti)",
    subtitle = "Acuti - Regime ordinario",
    caption = "Data Source: Ministero della Salute"
  )
p 


```


Table 2.2.3 shows the hospitalization rate for each region, by gender and year, for the activity *Acuti - Regime ordinario*.


```{r, echo=FALSE, include=TRUE}

HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime ordinario") %>%
  dplyr::group_by(ANNO)  %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), as.numeric) %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), round, digits = 3) %>% 
  dplyr::rename(`Tasso osp. Maschi` = "MASCHI",
                `Tasso osp. Femmine` = "FEMMINE") %>%
  dplyr::select(1:3, 5) %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 2.2.3"
  )

```


### TASSO PER TIPO DRG

The following section shows the evolution of the hospitalization rates per 1000 inhabitants for the activity *Acuti - Regime ordinario* by [**DRG**](#glossario) type. Indeed, the type of DRG could be *"C"*, that stands for **"CHIRURGICI"**, or *"M"*, that means **"MEDICI O NON CLASSIFICABILI"**. The distiction between the two is that the first DRG is characterised by the presence of surgery or other "significant" procedure, while the second one considers DRG with diagnostic or therapeutic procedures with no need of operating room.
As shown by the graph, the incidence of DRG type "M" is higher. For example, in 2019 there has been around 56 discharges per 1000 inhabitants for DRG with diagnostic or therapeutic procedures with no need of operating room(type "M"), while the incidence for the discharges characterized by DRG that required surgery(type "C") is around 44 discharges per 1000 inhabitants in the same year.


```{r,echo=FALSE, include=TRUE}
tipo_c <- SDO_ordinario %>%
  dplyr::group_by(ANNO) %>%
  dplyr::filter(TIPO == "C") %>%
  summarise(DIMISSIONI_TOT = sum(DIMISSIONI)) %>%
  pull(DIMISSIONI_TOT)

tipo_m <- SDO_ordinario %>%
  dplyr::group_by(ANNO) %>%
  dplyr::filter(TIPO == "M") %>%
  summarise(DIMISSIONI_TOT = sum(DIMISSIONI)) %>%
  pull(DIMISSIONI_TOT)

# tot_dim <- SDO_ordinario %>%
#   dplyr::group_by(ANNO) %>%
#   summarise(DIMISSIONI_TOT = sum(DIMISSIONI))

pop2019 <- 59816673
pop2018 <- 60483973
pop2017 <- 60589445
pop2016 <- 60665551

list_ita_pop <- c(pop2016, pop2017, pop2018, pop2019)


DRG_type = tibble(Tasso = c((tipo_c/list_ita_pop * 1000),
                            tipo_m/list_ita_pop * 1000), 
                  `TIPO DRG` = c(rep("C",4),rep("M", 4)),
                  Anno = rep(c("2016", "2017", "2018", "2019"),2))
   


DRG_type %>%
  ggplot(aes(x = Anno, y = Tasso, fill = `TIPO DRG`)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Tasso di ospedalizzazione per tipo DRG ( per 1000 abitanti)",
       subtitle = "Acuti - Regime ordinario",
       caption = "Data source : Ministero della Salute") + 
  xlab(" ") + 
  ylab("Tasso di osp. per 1000 abitanti")


```

The table below shows in detail the hospitalization rates previously discusssed and represented.

```{r, echo=FALSE, include=TRUE}

DRG_type %>%
  dplyr::mutate_at(vars(Tasso), round, digits = 4) %>%
  dplyr::rename(`Tasso di osp. per 1000 abitanti` = "Tasso") %>%
  DT::datatable(options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 2.2.4')
  



```




## Acuti - Regime diurno

Table 3.1.1 shows the hospitalization rates for the activity *Acuti - Regime diurno* for the 4 years of analysis and the corresponding percentage variation. An important decrease occurred in between 2016 and 2017 (-6.588%). The rate is decreasing over the 4 years of analysis.

```{r, echo=FALSE, include=TRUE}
HR_summary_act %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti -  diurno") %>%
  dplyr::mutate(`ATTIVITÀ` = "Acuti - Regime diurno") %>%  
  dplyr::rename(TASSO = "Tasso complessivo") %>%
  dplyr::mutate_at(vars(TASSO), round, digits = 3) %>%
  ungroup() %>%
  dplyr::mutate("VARIAZIONE %" = round((TASSO / lag(TASSO) -1) * 100, 
                                       digits   = 3)) %>%
  DT::datatable(
    options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 3.1.1')
  

```

### TASSO PER CATEGORIE DI ETÀ
The graph below shows that through the years of analysis the hospitalization rates by age category is decreased for all the age categories for the activity *Acuti - Regime diurno*.

```{r, echo=FALSE, include=TRUE}

HR_summary <- HR_age_cat %>%
  dplyr::filter(`Categoria età` != "TOTALE") %>%
  dplyr::select(`Categoria età`, `Acuti - Regime diurno`, ANNO) %>%
  dplyr::group_by(ANNO) 
  
HR_summary$`Categoria età` <- factor(HR_summary$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 14 anni",
                                            "Da 15 a 24 anni",
                                            "Da 25 a 44 anni", 
                                            "Da 45 a 64 anni", 
                                            "Da 65 a 74 anni", 
                                            "75 anni e oltre"))

  ggplot(data = HR_summary,
         aes(x = `Categoria età`, y = `Acuti - Regime diurno`, fill = ANNO)) +
  geom_bar( stat = "identity",
            position = "dodge") +
     theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (per età, per 1000 abitanti)",
              subtitle = "Acuti - Regime diurno",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") + 
        xlab("CATEGORIA ETÀ")



```


Table 3.1.2 represents the data about hospitalization rates for age category in *Acuti - Regime diurno* for each year.


```{r, echo=FALSE, include=TRUE}

HR_summary %>%
  dplyr::rename(`Tasso di ospedalizz. (per 1000 ab.)` = "Acuti - Regime diurno") %>%
  dplyr::mutate_at(vars(`Tasso di ospedalizz. (per 1000 ab.)`), round, digits = 3) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 3.1.2"
                )

  

```


### TASSO PER REGIONE E SESSO

The following graphs show the evolution of the hospitalization rate in Italy by region and gender in the period 2016-2019. The horizontal blue line displays the hospitalization rates in Italy for the male gender, the red one for the female gender.

```{r, echo=FALSE, include=TRUE}

gender_OR <- HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime diurno") %>%
  dplyr::group_by(ANNO) 

gender_OR$`REGIONE DI RESIDENZA` <- rep(c("Pie", "VdA", "Lom", "Blz", "Trn", "Ven",
                                      "FVG", "Lig", "EmR", "Tos", "Umb", "Mar",
                                      "Laz", "Abr", "Mol", "Cam", "Pug", "Bas",
                                      "Cal", "Sic", "Sar", "ITA"),4
                                      )
gender_OR$`REGIONE DI RESIDENZA` <- factor(gender_OR$`REGIONE DI RESIDENZA`,
                                          c( "Pie", "VdA", "Lom", "Blz",
                                             "Trn", "Ven", "FVG", "Lig", 
                                             "EmR", "Tos", "Umb", "Mar",      
                                             "Laz", "Abr", "Mol", "Cam", 
                                             "Pug", "Bas","Cal", "Sic",
                                             "Sar", "ITA")
                                          )
# reshape the dataframe in order to plot.

male <- gender_OR %>%
  dplyr::select(1, 2, 5) %>%
  dplyr::rename(TASSO = "MASCHI") %>%
  dplyr::mutate(GENERE = "MASCHI") 

female <- gender_OR %>%
  dplyr::select(1,3,5) %>%
  dplyr::rename(TASSO = "FEMMINE") %>%
  dplyr::mutate(GENERE = "FEMMINE") 


Gend_HR_total <- dplyr::bind_rows(male, female) %>%
  dplyr::mutate_at(vars(TASSO), as.numeric)

Italy_HR <- Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITA") %>%
  dplyr::rename(tasso_it = "TASSO")

Gend_HR_without_ita <- Gend_HR_total %>%
  filter(`REGIONE DI RESIDENZA` != "ITA")


p <- ggplot(data = Gend_HR_without_ita, aes( x = `REGIONE DI RESIDENZA`, 
                                 y = TASSO,
                                 fill = GENERE)) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "MASCHI"), aes(yintercept = tasso_it), 
             color = "blue") +
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "FEMMINE"), aes(yintercept = tasso_it), 
             color = "red") +
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +

  facet_wrap(~ANNO) +
  labs(
    title = "Tasso di ospedalizzazione per regione e sesso (per 1000 abitanti)",
    subtitle = "Acuti - Regime diurno",
    caption = "Data Source: Ministero della Salute"
  )
p 


```


Table 3.1.3 shows the hospitalization rate for each region, by gender and year, for the activity *Acuti - Regime diurno*.


```{r, echo=FALSE, include=TRUE}

HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime diurno") %>%
  dplyr::group_by(ANNO)  %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), as.numeric) %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), round, digits = 3) %>% 
  dplyr::rename(`Tasso osp. Maschi` = "MASCHI",
                `Tasso osp. Femmine` = "FEMMINE") %>%
  dplyr::select(1:3, 5) %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 3.1.3"
  )

```


### TASSO PER TIPO DRG

The following section shows the evolution of the hospitalization rates per 1000 inhabitants for the activity *Acuti - Regime diurno* by DRG type. 
In this activity, the incidence of DRG type "C" is higher. For example, in 2019 there has been around 17 discharges per 1000 inhabitants for DRG that required surgery (type "C") while dischages of DRG with diagnostic or therapeutic procedures with no need of operating room(type "M") have been around 12 per 1000 inhabitants in the same year.


```{r, echo=FALSE, include=TRUE}
tipo_c <- SDO_diurno %>%
  dplyr::group_by(ANNO) %>%
  dplyr::filter(TIPO == "C") %>%
  summarise(DIMISSIONI_TOT = sum(DIMISSIONI)) %>%
  pull(DIMISSIONI_TOT)

tipo_m <- SDO_diurno %>%
  dplyr::group_by(ANNO) %>%
  dplyr::filter(TIPO == "M") %>%
  summarise(DIMISSIONI_TOT = sum(DIMISSIONI)) %>%
  pull(DIMISSIONI_TOT)

# tot_dim <- SDO_ordinario %>%
#   dplyr::group_by(ANNO) %>%
#   summarise(DIMISSIONI_TOT = sum(DIMISSIONI))

pop2019 <- 59816673
pop2018 <- 60483973
pop2017 <- 60589445
pop2016 <- 60665551

list_ita_pop <- c(pop2016, pop2017, pop2018, pop2019)


DRG_type = tibble(Tasso = c((tipo_c/list_ita_pop * 1000),
                            tipo_m/list_ita_pop * 1000), 
                  `TIPO DRG` = c(rep("C",4),rep("M", 4)),
                  Anno = rep(c("2016", "2017", "2018", "2019"),2))
   


DRG_type %>%
  ggplot(aes(x = Anno, y = Tasso, fill = `TIPO DRG`)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Tasso di ospedalizzazione per tipo DRG ( per 1000 abitanti)",
       subtitle = "Acuti - Regime diurno",
       caption = "Data source : Ministero della Salute") + 
  xlab(" ") +
  ylab("Tasso di osp. per 1000 abitanti")


```

The table below shows in detail the hospitalization rates previously discusssed and represented.

```{r, echo=FALSE, include=TRUE}

DRG_type %>%
  dplyr::mutate_at(vars(Tasso), round, digits = 4) %>%
  dplyr::rename(`Tasso di osp. per 1000 abitanti` = "Tasso") %>%
  DT::datatable(options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 2.2.4')
  



```





## Riabilitazione - Regime ordinario

Table 4.1.1 shows the hospitalization rates for the activity *Riabilitazione - Regime ordinario* for the 4 years of analysis and the corresponding percentage variation. The hospitalization rate is decreasing over time.

```{r, echo=FALSE, include=TRUE}
HR_summary_act %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione -  ordinario") %>%
  dplyr::mutate(`ATTIVITÀ` = "Riabilitazione - Regime ordinario") %>%  
  dplyr::rename(TASSO = "Tasso complessivo") %>%
  dplyr::mutate_at(vars(TASSO), round, digits = 3) %>%
  ungroup() %>%
  dplyr::mutate("VARIAZIONE %" = round((TASSO / lag(TASSO) -1) * 100, 
                                       digits   = 3)) %>%
  DT::datatable(
    options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 4.1.1')
  

```

### TASSO PER CATEGORIE DI ETÀ
As shown by the graph, through the years of analysis the hospitalization rates remained stable for some of the age categories, while decreased for others, in particular for the categories: *"Da 65 a 74 anni"* and *"75 anni e oltre"* for the activity *Riabilitazione - Regime ordinario*.

```{r, echo=FALSE, include=TRUE}

HR_summary <- HR_age_cat %>%
  dplyr::filter(`Categoria età` != "TOTALE") %>%
  dplyr::select(`Categoria età`, `Riabilitazione - Regime ordinario`, ANNO) %>%
  dplyr::group_by(ANNO) 
  
HR_summary$`Categoria età` <- factor(HR_summary$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 14 anni",
                                            "Da 15 a 24 anni",
                                            "Da 25 a 44 anni", 
                                            "Da 45 a 64 anni", 
                                            "Da 65 a 74 anni", 
                                            "75 anni e oltre"))

  ggplot(data = HR_summary,
         aes(x = `Categoria età`, 
             y = `Riabilitazione - Regime ordinario`,
             fill = ANNO)) +
  geom_bar( stat = "identity",
            position = "dodge") +
     theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (per età, per 1000 abitanti)",
              subtitle = "Riabilitazione - Regime ordinario",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") + 
        xlab("CATEGORIA ETÀ")



```


Table 4.1.2 represents the data about hospitalization rates for age category in *Riabilitazione - Regime ordinario* for each year.


```{r, echo=FALSE, include=TRUE}

HR_summary %>%
  dplyr::rename(`Tasso di ospedalizz. (per 1000 ab.)` = "Riabilitazione - Regime ordinario") %>%
  dplyr::mutate_at(vars(`Tasso di ospedalizz. (per 1000 ab.)`), round, digits = 3) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 4.1.2"
                )

  

```


### TASSO PER REGIONE E SESSO

The following graphs show the evolution of the hospitalization rate in Italy by region and gender in the period 2016-2019. The horizontal blue line displays the hospitalization rates in Italy for the male gender, the red one for the female gender.

```{r, echo=FALSE, include=TRUE}

gender_OR <- HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione - Regime ordinario") %>%
  dplyr::group_by(ANNO) 

gender_OR$`REGIONE DI RESIDENZA` <- rep(c("Pie", "VdA", "Lom", "Blz", "Trn", "Ven",
                                      "FVG", "Lig", "EmR", "Tos", "Umb", "Mar",
                                      "Laz", "Abr", "Mol", "Cam", "Pug", "Bas",
                                      "Cal", "Sic", "Sar", "ITA"),4
                                      )
gender_OR$`REGIONE DI RESIDENZA` <- factor(gender_OR$`REGIONE DI RESIDENZA`,
                                          c( "Pie", "VdA", "Lom", "Blz",
                                             "Trn", "Ven", "FVG", "Lig", 
                                             "EmR", "Tos", "Umb", "Mar",      
                                             "Laz", "Abr", "Mol", "Cam", 
                                             "Pug", "Bas","Cal", "Sic",
                                             "Sar", "ITA")
                                          )
# reshape the dataframe in order to plot.

male <- gender_OR %>%
  dplyr::select(1, 2, 5) %>%
  dplyr::rename(TASSO = "MASCHI") %>%
  dplyr::mutate(GENERE = "MASCHI") 

female <- gender_OR %>%
  dplyr::select(1,3,5) %>%
  dplyr::rename(TASSO = "FEMMINE") %>%
  dplyr::mutate(GENERE = "FEMMINE") 


Gend_HR_total <- dplyr::bind_rows(male, female) %>%
  dplyr::mutate_at(vars(TASSO), as.numeric)

Italy_HR <- Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITA") %>%
  dplyr::rename(tasso_it = "TASSO")

Gend_HR_without_ita <- Gend_HR_total %>%
  filter(`REGIONE DI RESIDENZA` != "ITA")


p <- ggplot(data = Gend_HR_without_ita, aes( x = `REGIONE DI RESIDENZA`, 
                                 y = TASSO,
                                 fill = GENERE)) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "MASCHI"), aes(yintercept = tasso_it), 
             color = "blue") +
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "FEMMINE"), aes(yintercept = tasso_it), 
             color = "red") +
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +

  facet_wrap(~ANNO) +
  labs(
    title = "Tasso di ospedalizzazione per regione e sesso (per 1000 abitanti)",
    subtitle = "Riabilitazione - Regime ordinario",
    caption = "Data Source: Ministero della Salute"
  )
p 


```


Table 4.1.3 shows the hospitalization rate for each region, by gender and year, for the activity *Riabilitazione - Regime ordinario*.


```{r, echo=FALSE, include=TRUE}

HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione - Regime ordinario") %>%
  dplyr::group_by(ANNO)  %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), as.numeric) %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), round, digits = 3) %>% 
  dplyr::rename(`Tasso osp. Maschi` = "MASCHI",
                `Tasso osp. Femmine` = "FEMMINE") %>%
  dplyr::select(1:3, 5) %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 4.1.3"
  )

```


## Riabilitazione - Regime diurno

Table 5.1.1 shows the hospitalization rates for the activity *Riabilitazione - Regime diurno* for the 4 years of analysis and the corresponding percentage variation. An important reduction in the hospitalization rate is occurred in between 2017 an 2018 (-6.986%). The rate is decreasing until 2018, then there has been a little increase.

```{r, echo=FALSE, include=TRUE}
HR_summary_act %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione -  diurno") %>%
  dplyr::mutate(`ATTIVITÀ` = "Riabilitazione - Regime diurno") %>%  
  dplyr::rename(TASSO = "Tasso complessivo") %>%
  dplyr::mutate_at(vars(TASSO), round, digits = 3) %>%
  ungroup() %>%
  dplyr::mutate("VARIAZIONE %" = round((TASSO / lag(TASSO) -1) * 100, 
                                       digits   = 3)) %>%
  DT::datatable(
    options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 5.1.1')
  

```

### TASSO PER CATEGORIE DI ETÀ
As shown by the graph, through the years of analysis the hospitalization rates is increased for some of the age categories (*"Da 1 a 4 anni"* and *"Da 5 a 14 anni"*) and it is decreased for some others (*"Da 45 a 64 anni"*; *"Da 65 a 74 anni"* and *"75 anni e oltre"* ) for the activity *Riabilitazione - Regime diurno*.

```{r, echo=FALSE, include=TRUE}

HR_summary <- HR_age_cat %>%
  dplyr::filter(`Categoria età` != "TOTALE") %>%
  dplyr::select(`Categoria età`, `Riabilitazione - Regime diurno`, ANNO) %>%
  dplyr::group_by(ANNO) 
  
HR_summary$`Categoria età` <- factor(HR_summary$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 14 anni",
                                            "Da 15 a 24 anni",
                                            "Da 25 a 44 anni", 
                                            "Da 45 a 64 anni", 
                                            "Da 65 a 74 anni", 
                                            "75 anni e oltre"))

  ggplot(data = HR_summary,
         aes(x = `Categoria età`, 
             y = `Riabilitazione - Regime diurno`,
             fill = ANNO)) +
  geom_bar( stat = "identity",
            position = "dodge") +
     theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (per età, per 1000 abitanti)",
              subtitle = "Riabilitazione - Regime diurno",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") + 
        xlab("CATEGORIA ETÀ")



```


Table 5.1.2 represents the data about hospitalization rates for age category in *Riabilitazione - Regime diurno* for each year.


```{r, echo=FALSE, include=TRUE}

HR_summary %>%
  dplyr::rename(`Tasso di ospedalizz. (per 1000 ab.)` = "Riabilitazione - Regime diurno") %>%
  dplyr::mutate_at(vars(`Tasso di ospedalizz. (per 1000 ab.)`), round, digits = 3) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 5.1.2"
                )

  

```


### TASSO PER REGIONE E SESSO 

The following graphs show the evolution of the hospitalization rate in Italy by region and gender in the period 2016-2019. The horizontal blue line displays the hospitalization rates in Italy for the male gender, the red one for the female gender.
 
```{r, echo=FALSE, include=TRUE}

gender_OR <- HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione - Regime diurno") %>%
  dplyr::group_by(ANNO) 

gender_OR$`REGIONE DI RESIDENZA` <- rep(c("Pie", "VdA", "Lom", "Blz", "Trn", "Ven",
                                      "FVG", "Lig", "EmR", "Tos", "Umb", "Mar",
                                      "Laz", "Abr", "Mol", "Cam", "Pug", "Bas",
                                      "Cal", "Sic", "Sar", "ITA"),4
                                      )
gender_OR$`REGIONE DI RESIDENZA` <- factor(gender_OR$`REGIONE DI RESIDENZA`,
                                          c( "Pie", "VdA", "Lom", "Blz",
                                             "Trn", "Ven", "FVG", "Lig", 
                                             "EmR", "Tos", "Umb", "Mar",      
                                             "Laz", "Abr", "Mol", "Cam", 
                                             "Pug", "Bas","Cal", "Sic",
                                             "Sar", "ITA")
                                          )
# reshape the dataframe in order to plot.

male <- gender_OR %>%
  dplyr::select(1, 2, 5) %>%
  dplyr::rename(TASSO = "MASCHI") %>%
  dplyr::mutate(GENERE = "MASCHI") 

female <- gender_OR %>%
  dplyr::select(1,3,5) %>%
  dplyr::rename(TASSO = "FEMMINE") %>%
  dplyr::mutate(GENERE = "FEMMINE") 


Gend_HR_total <- dplyr::bind_rows(male, female) %>%
  dplyr::mutate_at(vars(TASSO), as.numeric)

Italy_HR <- Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITA") %>%
  dplyr::rename(tasso_it = "TASSO")

Gend_HR_without_ita <- Gend_HR_total %>%
  filter(`REGIONE DI RESIDENZA` != "ITA")


p <- ggplot(data = Gend_HR_without_ita, aes( x = `REGIONE DI RESIDENZA`, 
                                 y = TASSO,
                                 fill = GENERE)) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "MASCHI"), aes(yintercept = tasso_it), 
             color = "blue") +
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "FEMMINE"), aes(yintercept = tasso_it), 
             color = "red") +
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +

  facet_wrap(~ANNO) +
  labs(
    title = "Tasso di ospedalizzazione per regione e sesso (per 1000 abitanti)",
    subtitle = "Riabilitazione - Regime diurno",
    caption = "Data Source: Ministero della Salute"
  )
p 


```


Table 5.1.3 shows the hospitalization rate for each region, by gender and year, for the activity *Riabilitazione - Regime diurno*.


```{r, echo=FALSE, include=TRUE}

HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Riabilitazione - Regime diurno") %>%
  dplyr::group_by(ANNO)  %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), as.numeric) %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), round, digits = 3) %>% 
  dplyr::rename(`Tasso osp. Maschi` = "MASCHI",
                `Tasso osp. Femmine` = "FEMMINE") %>%
  dplyr::select(1:3, 5) %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 5.1.3"
  )

```



## Lungodegenza

Table 6.1.1 shows the hospitalization rates for the activity *Lungodegenza* for the 4 years of analysis and the corresponding percentage variation. A huge decrease occurred between 2016 and 2017 (-5.306%).

```{r, echo=FALSE, include=TRUE}
HR_summary_act %>%
  dplyr::filter(`ATTIVITÀ` == "Lungodegenza") %>%
  dplyr::mutate(`ATTIVITÀ` = "Lungodegenza") %>%  
  dplyr::rename(TASSO = "Tasso complessivo") %>%
  dplyr::mutate_at(vars(TASSO), round, digits = 3) %>%
  ungroup() %>%
  dplyr::mutate("VARIAZIONE %" = round((TASSO / lag(TASSO) -1) * 100, 
                                       digits   = 3)) %>%
  DT::datatable(
    options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 6.1.1')
  

```

### TASSO PER CATEGORIE DI ETÀ
The graph below shows that the hospitalization rate is almost 0 for the earlier age categories and is much higher for the last ones. Through the years of analysis the hospitalization rates decreased for these latter categories for the activity *Lungodegenza*.

```{r, echo=FALSE, include=TRUE}

HR_summary <- HR_age_cat %>%
  dplyr::filter(`Categoria età` != "TOTALE") %>%
  dplyr::select(`Categoria età`, `Lungodegenza`, ANNO) %>%
  dplyr::group_by(ANNO) 
  
HR_summary$`Categoria età` <- factor(HR_summary$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 14 anni",
                                            "Da 15 a 24 anni",
                                            "Da 25 a 44 anni", 
                                            "Da 45 a 64 anni", 
                                            "Da 65 a 74 anni", 
                                            "75 anni e oltre"))

  ggplot(data = HR_summary,
         aes(x = `Categoria età`, 
             y = `Lungodegenza`,
             fill = ANNO)) +
  geom_bar( stat = "identity",
            position = "dodge") +
     theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "TASSO DI OSPEDALIZZAZIONE (per età, per 1000 abitanti)",
              subtitle = "Lungodegenza",
       caption = "Data Source: Ministero della Salute") +
    ylab(" ") + 
        xlab("CATEGORIA ETÀ")



```


Table 6.1.2 represents the data about hopspitalization rates for age category in *Lungodegenza* for each year.


```{r, echo=FALSE, include=TRUE}

HR_summary %>%
  dplyr::rename(`Tasso di ospedalizz. (per 1000 ab.)` = "Lungodegenza") %>%
  dplyr::mutate_at(vars(`Tasso di ospedalizz. (per 1000 ab.)`), round, digits = 3) %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 6.1.2"
                )

  

```


### TASSO PER REGIONE E SESSO

The following graphs show the evolution of the hospitalization rate in Italy by region and gender in the period 2016-2019. The horizontal blue line displays the hospitalization rates in Italy for the male gender, the red one for the female gender.

```{r, echo=FALSE, include=TRUE}

gender_OR <- HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Lungodegenza") %>%
  dplyr::group_by(ANNO) 

gender_OR$`REGIONE DI RESIDENZA` <- rep(c("Pie", "VdA", "Lom", "Blz", "Trn", "Ven",
                                      "FVG", "Lig", "EmR", "Tos", "Umb", "Mar",
                                      "Laz", "Abr", "Mol", "Cam", "Pug", "Bas",
                                      "Cal", "Sic", "Sar", "ITA"),4
                                      )
gender_OR$`REGIONE DI RESIDENZA` <- factor(gender_OR$`REGIONE DI RESIDENZA`,
                                          c( "Pie", "VdA", "Lom", "Blz",
                                             "Trn", "Ven", "FVG", "Lig", 
                                             "EmR", "Tos", "Umb", "Mar",      
                                             "Laz", "Abr", "Mol", "Cam", 
                                             "Pug", "Bas","Cal", "Sic",
                                             "Sar", "ITA")
                                          )
# reshape the dataframe in order to plot.

male <- gender_OR %>%
  dplyr::select(1, 2, 5) %>%
  dplyr::rename(TASSO = "MASCHI") %>%
  dplyr::mutate(GENERE = "MASCHI") 

female <- gender_OR %>%
  dplyr::select(1,3,5) %>%
  dplyr::rename(TASSO = "FEMMINE") %>%
  dplyr::mutate(GENERE = "FEMMINE") 


Gend_HR_total <- dplyr::bind_rows(male, female) %>%
  dplyr::mutate_at(vars(TASSO), as.numeric)

Italy_HR <- Gend_HR_total %>%
  dplyr::filter(`REGIONE DI RESIDENZA` == "ITA") %>%
  dplyr::rename(tasso_it = "TASSO")

Gend_HR_without_ita <- Gend_HR_total %>%
  filter(`REGIONE DI RESIDENZA` != "ITA")


p <- ggplot(data = Gend_HR_without_ita, aes( x = `REGIONE DI RESIDENZA`, 
                                 y = TASSO,
                                 fill = GENERE)) + 
  geom_bar(stat = "identity",
           position = "dodge") + 
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "MASCHI"), aes(yintercept = tasso_it), 
             color = "blue") +
  geom_hline(data = Italy_HR %>% 
               filter(GENERE == "FEMMINE"), aes(yintercept = tasso_it), 
             color = "red") +
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +

  facet_wrap(~ANNO) +
  labs(
    title = "Tasso di ospedalizzazione per regione e sesso (per 1000 abitanti)",
    subtitle = "Lungodegenza",
    caption = "Data Source: Ministero della Salute"
  )
p 


```


Table 6.1.3 shows the hospitalization rate for each region, by gender and year, for the activity *Lungodegenza*.


```{r, echo=FALSE, include=TRUE}

HR_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Lungodegenza") %>%
  dplyr::group_by(ANNO)  %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), as.numeric) %>%
  dplyr::mutate_at(vars(MASCHI, FEMMINE), round, digits = 3) %>% 
  dplyr::rename(`Tasso osp. Maschi` = "MASCHI",
                `Tasso osp. Femmine` = "FEMMINE") %>%
  dplyr::select(1:3, 5) %>%
  
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 6.1.3"
  )

```


# CONFRONTO TASSI OSPEDALIZZAZIONE: ITALIA vs INGHILTERRA

This section gives an overview of the Italian and English population in 2019, then an analysis in England of the total number of surgeries (that can be compared to the total number of the italian **"dimissioni"**) and surgical incidences (the italian **"tasso di ospedalizzazione"**) in 2012 is made. Lastly, it shows a comparison between the overall hospitalization rates in Italy (2019) and England (2012).

## Popolazione: Italia vs Inghilterra

This first part shows the differences between English and Italian population in 2019, considering age category and gender. Data about population are provided by [**Istat**](http://dati.istat.it/Index.aspx?DataSetCode=DCIS_POPRES1) for Italy and [**UK Office for National Statistics**](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/bulletins/annualmidyearpopulationestimates/latest#ageing) for England. Note that for this latter, the population used in computing the surgical incidences is the one of 2012.


```{r, echo=FALSE,  include=TRUE, fig.width=10, fig.height=6}

pop_itaplot <- pop_ita2019 %>%
  dplyr::mutate(Country = "Italy") %>%
  dplyr::select(-c(stato_civile, Anno)) %>%
  dplyr::rename(popolazione = "Value") %>%
  dplyr::relocate(`Età`, popolazione, Sesso, `Categoria_età`)

eng_popplot <- eng_pop2019 %>%
  dplyr::mutate(Country = "England") %>%
  dplyr::mutate_at(vars(`Età`), as.character)

pop_to_plot <- dplyr::bind_rows(pop_itaplot, eng_popplot)

ggplot(data = pop_to_plot, 
       mapping = aes(x = ifelse(test = Sesso == "maschi", yes = -popolazione,
                                no = popolazione), 
                     y = `Categoria_età` , fill = Sesso)) +
  geom_col() +
  scale_x_symmetric(labels = abs) +  #from lemon package
  labs( title = "Piramide popolazione",
        subtitle = "Italia vs Inghilterra, 2019",
       x = "POPOLAZIONE",
       y = "ETÀ",
       caption = "Data Source: Istat"
       ) + 
  facet_wrap(~Country)


```



Table 3.1.1 summarises the data about the italian and english population in 2019, by age category and gender. The total population is 59.816.673, where 29.131.195 are male and 30.685.478 female.
In England the total population in 2019 is 56.286.961, where 27.827.831 are male and 28.459.130 female.

```{r, echo=FALSE, include=TRUE}

pop_to_plot %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 3.1.1')

```




## Numero di interventi e tassi di ospedalizzazione in Inghilterra per categoria di età.

The data collected for England about hospitalization rate consider two main activities, **In-patient surgeries** and **Out-patient surgeries**. The difference between the two is that the first one requires overnight hospitalization, while patients receiving out-patient care don’t need to spend a night in a hospital. These two activities can be compared respectively to the italian [**Ricovero Ordinario**](#glossario) and [**Day-hospital**](#glossario).


The following graph shows the total number of surgeries in England (In-patient and Out-patient) by age category and gender.
```{r, echo=FALSE, include=TRUE}

eng_surgery_M <- eng_surgery %>%
  dplyr::select(1,2) %>%
  dplyr::rename(`Number of Surgeries` = "MALE") %>%
  dplyr::mutate_at(vars(`Number of Surgeries`), as.numeric) %>%
  dplyr::mutate_at(vars(`Number of Surgeries`), round) %>%
  dplyr::mutate(GENDER = "Male")


eng_surgery_F <- eng_surgery %>%
  dplyr::select(1,3) %>%
  dplyr::rename(`Number of Surgeries` = "FEMALE") %>%
  dplyr::mutate_at(vars(`Number of Surgeries`), as.numeric) %>%
  dplyr::mutate_at(vars(`Number of Surgeries`), round) %>%
  dplyr::mutate(GENDER = "Female")

eng_sur_tot <- dplyr::bind_rows(eng_surgery_M, eng_surgery_F) %>%
   dplyr::filter(`AGE CATEGORY` != "Total") %>%
  dplyr::mutate_at(vars(`AGE CATEGORY`), factor)

eng_sur_tot$`AGE CATEGORY` <- factor(eng_sur_tot$`AGE CATEGORY`,
                                           c("0 - 4 Years", "5 - 9 Years",
                                           "10 - 14 Years", "15 - 19 Years", 
                                           "20 - 24 Years","25 - 29 Years",
                                           "30 - 34 Years", "35 - 39 Years",
                                           "40 - 44 Years", "45 - 49 Years", 
                                           "50 - 54 Years", "55 - 59 Years",
                                           "60 - 64 Years","65 - 69 Years", 
                                           "70 - 74 Years","75 - 79 Years",
                                           "80 - 84 Years", "85  &  over"))

eng_sur_tot %>%
  ggplot(aes(x = `AGE CATEGORY`, y = `Number of Surgeries`, fill = GENDER)) + 
  
  geom_bar(stat = "identity", position = "dodge") +
  
 theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "NUMBER OF SURGERIES (by age and gender)",
              subtitle = "England, 2019",
       caption = "Data Source: UK Office for National Statistics") +
    ylab(" ") + 
        xlab("AGE CATEGORY")


```


Table 3.1.2 summarises the data of the graph above. 

```{r, echo=FALSE, include=TRUE}

eng_sur_tot %>%
   DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 3.1.2"
  )


```





The following graph shows the the hospitalization rate in England for age category and gender, considering both In-patients and out-patients surgeries.

```{r, echo=FALSE, include=TRUE}

Eng_sur_inc_M <- England_sur_incidence %>%
  dplyr::select(1,2) %>%
  dplyr::rename(Incidence = "MALE") %>%
  dplyr::mutate(GENDER = "Male")

Eng_sur_inc_F <- England_sur_incidence %>%
  dplyr::select(1,3) %>%
  dplyr::rename(Incidence = "FEMALE") %>%
  dplyr::mutate(GENDER = "Female")


England_incidence <- dplyr::bind_rows(Eng_sur_inc_M, Eng_sur_inc_F) %>%
  dplyr::filter(`AGE CATEGORY` != "Total") %>%
  dplyr::mutate_at(vars(`AGE CATEGORY`), factor)

England_incidence$`AGE CATEGORY` <- factor(England_incidence$`AGE CATEGORY`,
                                           c("0 - 4 Years", "5 - 9 Years",
                                           "10 - 14 Years", "15 - 19 Years", 
                                           "20 - 24 Years","25 - 29 Years",
                                           "30 - 34 Years", "35 - 39 Years",
                                           "40 - 44 Years", "45 - 49 Years", 
                                           "50 - 54 Years", "55 - 59 Years",
                                           "60 - 64 Years","65 - 69 Years", 
                                           "70 - 74 Years","75 - 79 Years",
                                           "80 - 84 Years", "85  &  over"))






England_incidence %>%
  dplyr::mutate(Incidence = Incidence*1000) %>%
  ggplot(aes(x = `AGE CATEGORY`, y = Incidence, fill = GENDER)) + 
  
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "SURGICAL INCIDENCES (per 1000 inhabitants)",
              subtitle = "England, 2019",
       caption = "Data Source: UK Office for National Statistics") +
    ylab(" ") + 
        xlab("AGE CATEGORY")

```

Table 3.1.3 shows the data about surgical incidences represented in the graph above

```{r, echo=FALSE, include=TRUE}

England_incidence %>%
  dplyr::mutate(Incidence = Incidence*1000) %>%
  dplyr::mutate_at(vars(Incidence), round, digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 3.1.3"
  )

```


## Tassi di ospedalizzazione: Inghilterra vs Italia

The following graph provides an overview of the differences between the hospitalization rate in England (2012) and in Italy (2019) by age. The first thing noticed is that the female hospitalization rate is higher in both England and Italy.
In addition, the incidence is higher in England for both gender.


```{r, echo=FALSE, include=TRUE}

ita2019_M<- pop_ita2019  %>%
  filter(Sesso == "maschi") %>%
  pull(Value) %>%
  sum()

ita2019_F <- pop_ita2019  %>%
  filter(Sesso == "femmine") %>%
  pull(Value) %>%
  sum()

dimissioni_M_19 <- sdo_age_gender %>%
  filter(ANNO == "2019")  %>%
  pull(MASCHI) %>%
  sum()

dimissioni_F_19 <- sdo_age_gender %>%
  filter(ANNO == "2019")  %>%
  pull(FEMMINE) %>%
  sum()

incidence_England_tot <- England_sur_incidence %>%
  dplyr::filter(`AGE CATEGORY` == "Total")

incidence_it_en <- tibble(Incidence = c(dimissioni_M_19/ita2019_M * 1000,
                                        dimissioni_F_19/ita2019_F * 1000,
                                        incidence_England_tot$MALE * 1000,
                                        incidence_England_tot$FEMALE * 1000),
                          Gender = c("Male", "Female",
                                     "Male", "Female"),
                          Country = c("Italy", "Italy",
                                      "England", "England"))

incidence_it_en %>%
  ggplot(aes(x = Country, y = Incidence, fill = Gender)) +
geom_bar(stat = "identity", position = "dodge") + 
  labs(xlab = " ",
       title = "SURGICAL INCIDENCE, ITALY vs ENGLAND")




```
 
Table 3.1.4 shows the data about hospitalization rate represented in the bar chart above.

```{r, echo=FALSE, include=TRUE}

incidence_it_en %>%
  dplyr::mutate_at(vars(Incidence), round, digits = 3) %>%
   DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 3.1.4"
  )



```



# CONFRONTO DATI DIMISSIONI OSPEDALIERE: ISTAT VS MINISTERO DELLA SALUTE

The following section provides a comparison between the data about hospitalization discharges in 2019 collected from two different sources: Istat and Italian Ministry of Health.

Istat provided data just for the activity *Acuti*, distinguishing between *"Ordinario"* and *"Day-Hospital"*, that are respectively given by *Acuti - Regime ordinario* and *Acuti - Regime diurno* from Ministry of Health.

```{r, echo=FALSE, include=TRUE}

istat_tot_dism <- istat_discharge %>% 
  dplyr::filter(`REGIME DI RICOVERO` != "totale",
         GENERE == "totale",
         `AREA DI CITTADINANZA` == "Mondo",
         `CLASSE DI ETÀ` != "totale") %>%
  dplyr::group_by(`REGIME DI RICOVERO`) %>%
  dplyr::rename(`attività` = `REGIME DI RICOVERO`) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(`attività` = c("Acuti - Regime diurno",
                                         "Acuti - Regime ordinario")) %>%
  dplyr::mutate(Source = "Istat")

sum_ordinario_19 <- summary_ordinario %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(`attività` = "Acuti - Regime ordinario") %>%
  dplyr::select(-`VARIAZIONE %`)

sum_diurno_19 <- summary_diurno %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::select(-(3)) %>% 
  mutate(`attività` = "Acuti - Regime diurno")

summary_tot <- dplyr::bind_rows(sum_ordinario_19, sum_diurno_19) %>%
  dplyr::select(-ANNO) %>%
  dplyr::mutate(Source = "Ministero della Salute") %>%
  dplyr::relocate(`attività`, DIMISSIONI, Source)

summary_plotted <- dplyr::bind_rows(summary_tot, istat_tot_dism) %>%
  dplyr::mutate(ANNO = "2019")

summary_plotted %>%
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 4.1.1')

```


As shown by the table above, there is a slight difference in the data provided by the two sources for both *Regime ordinario* and *Regime diurno*.


## Istat, Acuti - Regime ordinario

This part focuses on the analysis of hospitalization discharges for the activity *"Acuti - Regime ordinario"* considering data provided by istat.
In the data provided by istat, the age categories are a very limited number compared to the ones of Italian Ministry. This is shown in the graph below.

```{r, echo=FALSE, include=TRUE}

istat_byage <- istat_discharge %>% 
  filter(`REGIME DI RICOVERO` == "ordinario",
         GENERE == "totale",
         `AREA DI CITTADINANZA` == "Mondo",
         `CLASSE DI ETÀ` != "totale" ) %>%
  dplyr::mutate_at(vars(`CLASSE DI ETÀ`), factor)

istat_byage$`CLASSE DI ETÀ` <- factor(istat_byage$`CLASSE DI ETÀ`,
                                          c("fino a 17 anni",
                                            "18-24 anni",
                                            "25-44 anni",
                                            "45-64 anni",
                                            "65 anni e più",
                                            "non indicato"
                                            ))
istat_byage %>%
 
  ggplot(., aes(x = `CLASSE DI ETÀ`,y = DIMISSIONI, fill = `CLASSE DI ETÀ` )) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ (Acuti - Regime ordinario)",
       subtitle = "Italia, 2019",
       caption = "Data Source: Istat")

  


```


Table 4.1.2 provides a summary of the hospitalization discharges by age category in 2019 in Italy. 


```{r, echo=FALSE, include=TRUE}

istat_byage %>%
  dplyr::select(-c(2,4,5)) %>%
  dplyr::mutate(`REGIME DI RICOVERO` = "Acuti - Regime ordinario") %>%
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 4.1.2')


```



Data provided by Istat consider the distinction between hospitalization discharges by gender, as well as data provided by Italian Ministry of Health.
As shown by the graph below, the numbers are slightly different due to the fact that the Italian Ministry data, in distinguishing between male and female, does not consider discharges where age or gender were not well specified.


```{r, echo=FALSE, include=TRUE}

istat_gender <- istat_discharge %>% 
  filter(`REGIME DI RICOVERO` == "ordinario",
         GENERE != "totale",
         `AREA DI CITTADINANZA` == "Mondo",
         `CLASSE DI ETÀ` != "totale" & `CLASSE DI ETÀ` != "non indicato") %>%
  group_by(GENERE) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(Source = "Istat")
   

MM <- sdo_age_gender %>%
  filter(`ATTIVITÀ` == "Acuti - Regime ordinario",
         ANNO == "2019") %>%
  pull(MASCHI) %>%
  sum()

FF <- sdo_age_gender %>%
  filter(`ATTIVITÀ` == "Acuti - Regime ordinario",
         ANNO == "2019") %>%
  pull(FEMMINE) %>%
  sum()

sdo_gen <- tibble(GENERE = c("maschi", "femmine"), DIMISSIONI = c(MM,FF)) %>%
  dplyr::mutate(Source = "Ministero della Salute")


gender_plot <- dplyr::bind_rows(sdo_gen, istat_gender)

gender_plot %>%
  ggplot( aes(x = GENERE, 
             y = DIMISSIONI,
             fill = Source)) +
  geom_bar( stat = "identity",
            position = "dodge") + 
  labs(title = "DIMISSIONI PER GENERE",
       subtitle = "Istat vs Ministero della Salute, Italia 2019")


```

An important difference in data provided by istat is the distinction between country of origin of the patient, as shown in the graph below.

```{r, echo=FALSE, include=TRUE}

istat_country<- istat_discharge %>% 
  dplyr::filter(`REGIME DI RICOVERO` == "ordinario",
         GENERE == "totale",
         `AREA DI CITTADINANZA` %in% c("Europa", "Asia", "Africa", "America", 
                                       "Oceania", "Apolide", "Unknown")) %>%
  dplyr::filter(`CLASSE DI ETÀ` != "totale" 
                & `CLASSE DI ETÀ` != "non indicato") %>%
  dplyr::mutate(`REGIME DI RICOVERO` = "Acuti - Regime ordinario")

istat_country$`CLASSE DI ETÀ` <- factor(istat_country$`CLASSE DI ETÀ`,
                                          c("fino a 17 anni",
                                            "18-24 anni",
                                            "25-44 anni",
                                            "45-64 anni",
                                            "65 anni e più"
                                            ))

p <- ggplot(istat_country,
            aes(x = `CLASSE DI ETÀ`,
                y = DIMISSIONI,
                fill = `AREA DI CITTADINANZA`)) + 
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "DIMISSIONI REGIME ORDINARIO PER AREA ED ETÀ",
  subtitle = "Italia, 2019",
  caption = "Data Source: Istat")
p
```


The pie chart below shows the geographical area of the patient without discriminate by age category. As shown, the discharges of people from *Oceania* and people that are stateless (*Apolide*) are negligible. The most important proportion of the table is, of course, by patients of Europe. This is due to the fact that the number of discharges of Italian people is the highest, namely 5.615.223.

```{r, echo=FALSE, include=TRUE}

istat_country_pie<- istat_discharge %>% 
  dplyr::filter(`REGIME DI RICOVERO` == "ordinario",
         GENERE == "totale",
         `AREA DI CITTADINANZA` %in% c("Europa", "Asia", "Africa", "America", 
                                       "Oceania", "Apolide", "Unknown")) %>%
  dplyr::filter(`CLASSE DI ETÀ` == "totale" 
                & `CLASSE DI ETÀ` != "non indicato") %>%
  dplyr::mutate(`REGIME DI RICOVERO` = "Acuti - Regime ordinario") %>%
  dplyr::mutate(PERCENTUALE = DIMISSIONI / sum(DIMISSIONI))

p <- ggplot(istat_country_pie, aes(x = " ", y = PERCENTUALE,
                                   fill = `AREA DI CITTADINANZA`)) + 
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0)
p + labs(caption = "Data Source: Istat") + 
  xlab("")+
  ylab("")
```




The following table summarises the data shown in the graphs above.
```{r, echo=FALSE, include=TRUE}
istat_country %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 4.1.3"
  )

```

## Istat, Acuti - Regime diurno
This second part is concerned to the analysis of hospitalization discharges for the activity *"Acuti - Regime diurno"* considering data provided by istat.

This first graph shows the hospitalization discharges taking into account the different age category.

```{r, echo=FALSE, include=TRUE}

istat_byage_d <- istat_discharge %>% 
  filter(`REGIME DI RICOVERO` == "day-hospital",
         GENERE == "totale",
         `AREA DI CITTADINANZA` == "Mondo",
         `CLASSE DI ETÀ` != "totale" ) %>%
  dplyr::mutate_at(vars(`CLASSE DI ETÀ`), factor)

istat_byage_d$`CLASSE DI ETÀ` <- factor(istat_byage_d$`CLASSE DI ETÀ`,
                                          c("fino a 17 anni",
                                            "18-24 anni",
                                            "25-44 anni",
                                            "45-64 anni",
                                            "65 anni e più",
                                            "non indicato"
                                            ))
istat_byage_d %>%
 
  ggplot(., aes(x = `CLASSE DI ETÀ`,y = DIMISSIONI, fill = `CLASSE DI ETÀ` )) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "DIMISSIONI PER ETÀ (Acuti - Regime diurno)",
       subtitle = "Italia, 2019",
       caption = "Data Source: Istat")

  


```

Table 4.1.4 provides a summary of the hospitalization discharges by age category in 2019 in Italy. 

```{r, echo=FALSE, include=TRUE}

istat_byage_d %>%
  dplyr::select(-c(2,4,5)) %>%
  dplyr::mutate(`REGIME DI RICOVERO` = "Acuti - Regime diurno") %>%
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 4.1.4')


```

Data provided by Istat consider the distinction between hospitalization discharges by gender, as well as data provided by Italian Ministry of Health.
Even in the case of *Acuti - Regime diurno* the numbers are slightly different due to the fact that the Italian Ministry data, in distinguishing between male and female, does not consider discharges where age or gender were not well specified.


```{r, echo=FALSE, include=TRUE}

istat_gender_d <- istat_discharge %>% 
  filter(`REGIME DI RICOVERO` == "day-hospital",
         GENERE != "totale",
         `AREA DI CITTADINANZA` == "Mondo",
         `CLASSE DI ETÀ` != "totale" & `CLASSE DI ETÀ` != "non indicato") %>%
  group_by(GENERE) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(Source = "Istat")
   

MM <- sdo_age_gender %>%
  filter(`ATTIVITÀ` == "Acuti - Regime diurno",
         ANNO == "2019") %>%
  pull(MASCHI) %>%
  sum()

FF <- sdo_age_gender %>%
  filter(`ATTIVITÀ` == "Acuti - Regime diurno",
         ANNO == "2019") %>%
  pull(FEMMINE) %>%
  sum()

sdo_gen <- tibble(GENERE = c("maschi", "femmine"), DIMISSIONI = c(MM,FF)) %>%
  dplyr::mutate(Source = "Ministero della Salute")


gender_plot <- dplyr::bind_rows(sdo_gen, istat_gender_d)

gender_plot %>%
  ggplot( aes(x = GENERE, 
             y = DIMISSIONI,
             fill = Source)) +
  geom_bar( stat = "identity",
            position = "dodge") + 
  labs(title = "DIMISSIONI PER GENERE",
       subtitle = "Istat vs Ministero della Salute, Italia 2019")


```

An important difference in data provided by istat is the distinction between country of origin of the patient, as shown in the graph below.

```{r, echo=FALSE, include=TRUE}

istat_country_d<- istat_discharge %>% 
  dplyr::filter(`REGIME DI RICOVERO` == "day-hospital",
         GENERE == "totale",
         `AREA DI CITTADINANZA` %in% c("Europa", "Asia", "Africa", "America", 
                                       "Oceania", "Apolide", "Unknown")) %>%
  dplyr::filter(`CLASSE DI ETÀ` != "totale" 
                & `CLASSE DI ETÀ` != "non indicato") %>%
  dplyr::mutate(`REGIME DI RICOVERO` = "Acuti - Regime diurno")

istat_country_d$`CLASSE DI ETÀ` <- factor(istat_country_d$`CLASSE DI ETÀ`,
                                          c("fino a 17 anni",
                                            "18-24 anni",
                                            "25-44 anni",
                                            "45-64 anni",
                                            "65 anni e più"
                                            ))

p <- ggplot(istat_country_d,
            aes(x = `CLASSE DI ETÀ`,
                y = DIMISSIONI,
                fill = `AREA DI CITTADINANZA`)) + 
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "DIMISSIONI REGIME DIURNO, PER AREA ED ETÀ",
  subtitle = "Italia, 2019",
  caption = "Data Source: Istat")
p
```


The pie chart below shows the geographical area of the patient without discriminate by age category for *Acuti - Regime diurno*. Even in this case, the discharges of people from *Oceania* and people that are stateless (*Apolide*) are negligible. The most important proportion of the table is, of course, by patients of Europe. This is due to the fact that the number of discharges of Italian people is the highest, that is 1.647.029.

```{r, echo=FALSE, include=TRUE}

istat_country_pie_d<- istat_discharge %>% 
  dplyr::filter(`REGIME DI RICOVERO` == "day-hospital",
         GENERE == "totale",
         `AREA DI CITTADINANZA` %in% c("Europa", "Asia", "Africa", "America", 
                                       "Oceania", "Apolide", "Unknown")) %>%
  dplyr::filter(`CLASSE DI ETÀ` == "totale" 
                & `CLASSE DI ETÀ` != "non indicato") %>%
  dplyr::mutate(`REGIME DI RICOVERO` = "Acuti - Regime diurno") %>%
  dplyr::mutate(PERCENTUALE = DIMISSIONI / sum(DIMISSIONI))

p <- ggplot(istat_country_pie_d, aes(x = " ", y = PERCENTUALE,
                                   fill = `AREA DI CITTADINANZA`)) + 
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0)
p + labs( caption = "Data Source: Istat") + 
  xlab("") +
  ylab("")
```




The following table summarises the data shown in the graphs above.
```{r, echo=FALSE, include=TRUE}
istat_country_d %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 4.1.5"
  )

```




## Tasso di ospedalizzazione per l'attività Acuti: dati Istat vs Ministero della Salute

The following paragraph focuses on the differences in the overall hospitalization rate considering the two different data sources (Istat and Ministero della Salute) in Italy in 2019 for the activity *Acuti*, so, considering the sum of discharges in *Regime ordinario* and *Regime diurno*.


This table shows the hospitalization rates per 1000 inhabitants(Italy, 2019), considering the two different data sources, for the activity *Acuti* (taking into account of the discharges in *Regime ordinario* and *Regime diurno*). As shown, the difference is particularly small. Indeed, the number of disharges is almost the same in considering the two sources.

```{r, echo=FALSE, include=TRUE}
# consider the sum of discharges for ACUTI for ordinario and diurno (SDO data) 

SDO_dim_acuti <- summary_tab %>%
  dplyr::filter(ANNO == "2019",
                `ATTIVITÀ` == "Acuti - Regime ordinario" |
                  `ATTIVITÀ` == "Acuti - Regime diurno") %>%
  dplyr::pull(DIMISSIONI) %>%
  sum()

istat_dim_acuti <- istat_discharge %>%
  dplyr::filter(GENERE == "totale",
                `AREA DI CITTADINANZA` == "Mondo",
                `REGIME DI RICOVERO` != "totale",
                `CLASSE DI ETÀ` != "totale") %>%
  dplyr::pull(DIMISSIONI) %>%
  sum()



comp_table <- dplyr::tibble(Source = c("Istat", "Ministero della Salute"),
                            Dimissioni_Acuti = c((SDO_dim_acuti/sum(pop_ita2019$Value)) 
                                                 * 1000,
                                                 (istat_dim_acuti/sum(pop_ita2019$Value))
                                                 * 1000)) 
comp_table %>% 
  dplyr::mutate_at(vars(Dimissioni_Acuti), round, digits = 4) %>%
  dplyr::rename("Tasso osp. per 1000 abitanti" = "Dimissioni_Acuti") %>%
  dplyr::mutate("Numero di dimissioni" = c(istat_dim_acuti, SDO_dim_acuti)) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 4.1.6"
  )
  



```












\pagebreak

# GLOSSARIO {-}

**Accessi**: presence in hospital without overnight stay, for the provision of day-hospital benefits.

**Attività**: type of hospital care; it includes: *ricoveri per acuti*, *riabilitazione* and *lungodegenza*.


**Day Hospital (diurno)**: characterized by: programmed hospitalization; it is limited to part of the day, it does not include overnight staying; it provides multiprofessional or plurispecialistic services that require a running time that deviates sharply from that necessary for a normal outpatient performance.

**DRG – DIAGNOSIS RELATED GROUP**: 
category of hospital admissions defined in such a way that they have similar clinical characteristics and require homogeneous volumes of hospital resources for their treatment. The following information is required to assign each admission episode to a specific DRG: the main diagnosis of discharge, all secondary diagnoses, all surgical procedures and the main diagnostic and therapeutic procedures, age, sex and mode of discharge. The attribution is carried out using an algorithm that analyzes the above information and determines the group to which it belongs.

**Giorno di Degenza**: period of 24 hours during which a bed is occupied.  

**MDC - MAJOR DIAGNOSTIC CATEGORY (MDC)**:
the DRG system provides a higher level of grouping based on a criterion of clinical relevance (anatomical or etiological): based on the main diagnosis reported in the *Scheda di Dimissione Ospedaliera*, the MDC category is assigned; subsequently, a specific DRG is assigned among those belonging to the selected MDC. 

**Ricoveri di Lungodegenza**: these are the admissions of patients discharged from departments belonging to the *LUNGODEGENTI* discipline.

**Regime di ricovero**: distinguishes between “RICOVERO ORDINARIO”, that requires for overnight stay in the hospital, and “RICOVERO IN DAY HOSPITAL”,characterized by the presence in hospital just for a part of the day.

**Ricoveri per Acuti**: all cases of discharges from departments other than those classified as rehabilitative or long-term; healthy infants are also excluded. 

**Ricoveri di Riabilitazione**: identified as admissions of patients discharged from departments belonging to the disciplines: *UNITA’ SPINALE*, *RECUPERO E RIABILITAZIONE FUNZIONALE*, *NEURORIABILITAZIONE*, and hospitalization in rehabilitation institutions only (LUNGODEGENZA excluded). 

**Ricovero Ordinario**: provides for the allocation of a bed and the continuous stay of the patient with overnight stay in the hospital.


**Tasso di ospedalizzazione**: the hospitalization rate provides a
summary measure on the use of hospitalization of the population under study. It is calculated as the ratio of the number of hospitalizations, everywhere
carried out, relating to persons resident in a given region, and the total population resident there.

**Valore Soglia**: value of stay duration, specific for each DRG, above or below which a stay in *Regime ordinario* is considered "abnormal for the duration of the stay".