---
title: "SDO analysis"
author: "Marco Gaudio"
date: "2/3/2021"
output:
  html_document:
    code_folding: show
    highlight: tango
    keep_md: no
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  github_document:
    toc: yes
---

```{r, include=FALSE}
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE, 
                      include = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = FALSE)
library(tidymodels)
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(DT)
```


The aim of this document is to analyze the datas related to the italian *Schede Dimissioni Ospedaliere* (SDO), in the period 2016 - 2019. 
The data are provided by the [**Italian Ministry of Health**](http://www.salute.gov.it/portale/temi/p2_6.jsp?lingua=italiano&id=1237&area=ricoveriOspedalieri&menu=vuotolink).

The document is composed by 3 sections: 

* **section 1** provides an analysis for each of the 5 activities (Acuti Regime ordinario;       Acuti Regime diurno, Riabilitazione Regime ordinario; Riabilitazione Regime diurno;        Lungodegenza) and considering the differences in the period studied. 

* **section 2** analyses the hospitalization rates in Italy in the period 2016-2019, considering different region, gender and age category.

* **section 3** provides a comparison between  hospitalization rates in Italy and hospitalization rates in UK.


```{r}
# construct a summary table
summary_ordinario <- matching_tables %>% 
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Acuti - Regime ordinario")

summary_diurno <- matching_tables_2.2.11 %>% 
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Acuti - Regime diurno")

summary_riab_ord <- SDO_ordinario_riab %>%
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Riabilitazione - Regime ordinario")

summary_riab_diurno <- SDO_diurno_riab %>%
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Riabilitazione - Regime diurno")

summary_lungodegenza <- SDO_lungodegenza %>%
  group_by(ANNO) %>%
  summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3)) %>%
  mutate(`ATTIVITÀ` = "Riabilitazione - Regime diurno")

summary_tab <- dplyr::bind_rows(summary_ordinario,
                                summary_diurno,
                                summary_riab_ord,
                                summary_riab_diurno,
                                summary_lungodegenza)

```

The following graph shows the total number of hospital discharges divided by activity for the 4 years of analysis.

```{r, echo=FALSE, include=TRUE}

  p <- ggplot(summary_tab, aes(x = ANNO, y = DIMISSIONI, fill = `ATTIVITÀ`)) + 
  geom_bar(stat = "identity") +
  labs(title = 'TOTAL DISCHARGES IN ITALY') 
p
```

Table 1. provides the total number of discharges for all the activities and its percentage variation through the period of analysis. As shown here, the total number of hospital discharges is decreased from 2016 to 2019. In particular, from 2018 to 2019 the number has dropped from 8.522.456 to 8.356.874, with a decrease of 1.733%.

```{r, echo=FALSE, include=TRUE}

summary_tab %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100,
                                        digits = 3)) %>%
DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.')

```



# ATTIVITÀ

## Acuti - Regime ordinario.
the following section is aimed to analyze the discharge in *Acuti - Regime ordinario*, from 2016 to 2019 in Italy.

### DIMISSIONI

Table 1.1 provides a summary of the hospedalization discharges in *"Acuti - Regime ordinario"* and the variation from year to year.

```{r, echo=FALSE, include=TRUE}
summary_ordinario <- matching_tables %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_ordinario %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 1.1')



```

This bar chart shows the hospital discharges by [MDC](#glossario) for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_ordinario %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "HOSPITAL DISCHARGES BY MDC",
              subtitle = "Acuti - Regime ordinario")

  
```

As shown from the bar chart, over the 4 years of analysis, the MDC class with the greatest number of dicharges for the activity *Acuti - Regime ordinario* is the **MDC 05 - Malattie e disturbi dell'apparato cardiocircolatorio**, followed by **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo** and **MDC 04 - Malattie e disturbi dell'apparato respiratorio**.  



The following table provides a summary of the discharges for each MDC class and for each year.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime ordinario for the 4 years
overall_ordinario <- matching_tables %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_ordinario %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.2')
```

### PERCENTUALE GIORNI DIMISSIONI PER MDC.

Days before discharge in 2019 for MDC in percentage.


```{r, echo=FALSE, include=TRUE}

percentage_tab <- function(year){
  perc <- SDO_ordinario %>%
  filter(ANNO == year) %>%
  mutate(percentage_dim1gg = `% DIMISSIONI DI 1 GIORNO`/100,
         percentage_dim2gg = `% DIMISSIONI DI 2-3 GIORNI`/100,
         percentage_dim4gg = `% DIMISSIONI TRA 4 GIORNI E VALORE SOGLIA`/100,
         percentage_dimooltre = `% DIMISSIONI OLTRE VALORE SOGLIA`/100) 
# compute the percentage for the whole MDC

dimissioni <- as.matrix(perc$DIMISSIONI, ncol = 1)  
percent_oneday <-  as.matrix(perc$percentage_dim1gg, nrow = 1) 
percent_twoday <- as.matrix(perc$percentage_dim2gg, nrow = 1)
percent_fourday <- as.matrix(perc$percentage_dim4gg, nrow = 1)
percentbeyond <- as.matrix(perc$percentage_dimooltre, nrow = 1)

dimissioni_1 <- round((t(percent_oneday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_2 <- round((t(percent_twoday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_4 <- round((t(percent_fourday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_oltre <- round((t(percentbeyond) %*% dimissioni )/ sum(perc$DIMISSIONI), 
                          digits = 3)
return(c(dimissioni_1,dimissioni_2, dimissioni_4, dimissioni_oltre))
}
perc <- SDO_ordinario %>%
  filter(ANNO == "2019") %>%
  mutate(percentage_dim1gg = `% DIMISSIONI DI 1 GIORNO`/100,
         percentage_dim2gg = `% DIMISSIONI DI 2-3 GIORNI`/100,
         percentage_dim4gg = `% DIMISSIONI TRA 4 GIORNI E VALORE SOGLIA`/100,
         percentage_dimooltre = `% DIMISSIONI OLTRE VALORE SOGLIA`/100) 
# compute the percentage for the whole MDC

dimissioni <- as.matrix(perc$DIMISSIONI, ncol = 1)  
percent_oneday <-  as.matrix(perc$percentage_dim1gg, nrow = 1) 
percent_twoday <- as.matrix(perc$percentage_dim2gg, nrow = 1)
percent_fourday <- as.matrix(perc$percentage_dim4gg, nrow = 1)
percentbeyond <- as.matrix(perc$percentage_dimooltre, nrow = 1)

dimissioni_1 <- round((t(percent_oneday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_2 <- round((t(percent_twoday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_4 <- round((t(percent_fourday) %*% dimissioni )/ sum(perc$DIMISSIONI), digits = 3)
dimissioni_oltre <- round((t(percentbeyond) %*% dimissioni )/ sum(perc$DIMISSIONI), 
                          digits = 3)

#PROVA
# degenza_tot_gg <- SDO_ordinario %>%
#   dplyr::filter(ANNO == "2019",
#          CLASSE_MDC == "MDC 01 - Malattie e disturbi del sistema nervoso") %>%
#   dplyr::summarise(sum(DIMISSIONI)) %>%
#   dplyr::pull() %>%
#   as.numeric()
# 
# dimissioni_1*degenza_tot_gg+
# dimissioni_2*degenza_tot_gg+
# dimissioni_4*degenza_tot_gg+
# dimissioni_oltre*degenza_tot_gg

mycols <- c("% DIMISSIONI DI 1 GIORNO",                 
            "% DIMISSIONI DI 2-3 GIORNI",              
            "% DIMISSIONI TRA 4 GIORNI E VALORE SOGLIA",
            "% DIMISSIONI OLTRE VALORE SOGLIA")

prop_table <- function(year){
  discharges_prop <- percentage_tab(year)
tabella_plot <- tidyr::tibble(class = mycols,
                       proportion = discharges_prop) %>%
  dplyr::arrange(-proportion) %>%
  dplyr::mutate(Per_cumsum = cumsum(proportion)) %>%
  dplyr::mutate(ANNO = year)

}
Years <- c("2016", "2017", "2018", "2019")

summary_list <- purrr::map(Years, prop_table)

summary_prop <- dplyr::bind_rows(summary_list) 

discharges_2019 <- percentage_tab("2019")
tabella_plot_2019 <- tibble(class = mycols,
                       proportion = discharges_2019) %>%
  arrange(-proportion) %>%
  mutate(Per_cumsum = cumsum(proportion))

# ggplot(tabella_plot, aes(x = "", y = proportion, fill = class)) +
#   geom_bar(width = 1, stat = "identity", color = "white") +
#   geom_col() +
#   geom_text(aes(y = Per_cumsum - proportion/2, label = proportion*100), color = "white")+ 
#   coord_polar("y", start = 0)+
#   theme_void()

ggplot(data = summary_prop, aes(x=1, y=proportion, fill=class)) +
  geom_col() +
  geom_text(aes(label = proportion*100), position = position_stack(vjust = 0.5))+
  coord_polar(theta = "y") + 
  theme_void() +
  labs(title = "% DIMISSIONI PER GIORNO",
       fill = "") +
  facet_wrap(~ANNO)



```






The bar chart below shows the distribution of discharges in *Acuti - Regime ordinario* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime ordinario") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5))

  
  




```




### GIORNATE DI DEGENZA {.tabset}

The following graph shows the [*Giorni di Degenza*](#glossario) for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_ordinario %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(GIORNI_DEGENZA = sum(`GIORNI DI DEGENZA`)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = GIORNI_DEGENZA,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "GIORNI DI DEGENZA",
              subtitle = "Acuti - Regime ordinario")


```


Table 1.3 provides a summary for *Giorni di Degenza* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_ordinario_degenza <- matching_tables %>%
  dplyr::select(CLASSE_MDC, `GIORNATE DI DEGENZA`, ANNO)

overall_ordinario_degenza %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 1.3"
                )

```

The following graph shows the Mean value for *Giorni di Degenza* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```

#### 2017 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}
sum_deg <- SDO_ordinario %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3))

ggplot(data = sum_deg, mapping = aes(x = forcats::fct_reorder(CLASSE_MDC,
                                                             MEAN_DEGENZA), 
                                     y = MEAN_DEGENZA)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Acuti - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  


```

### {-}

Table 1.3 provides a summary of Max, Min and Mean values for *"Giorni di Degenza"* for each MDC.
```{r, echo=FALSE, include=TRUE}

SDO_ordinario %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_DEGENZA = max(`Degenza massima`),
            MIN_DEGENZA = min(`Degenza minima`),
            MEAN_DEGENZA = round(mean(`DEGENZA MEDIA (giorni)`), digits = 3)) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 1.3'
    
  )

```



## Acuti - Regime diurno.

### DIMISSIONI

Table 2.1 provides a summary of the hospedalization discharges in *"Acuti - Regime diurno"* and the variation from year to year.


```{r, echo=FALSE, include=TRUE}
summary_diurno <- matching_tables_2.2.11 %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_diurno %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 2.1')



```


This bar chart shows the hospital discharges by MDC for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_diurno %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "HOSPITAL DISCHARGES BY MDC",
              subtitle = "Acuti - Regime diurno")

  
```


In the case of *Acuti - Regime diurno*, over the 4 years of analysis, the MDC class with the greatest number of dicharges is the **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**, followed by **MDC 17 - Malattie e disturbi mieloproliferativi e neoplasie scarsamente differenziate** and **MDC 13 -Malattie e disturbi dell'apparato riproduttivo femminile**.  


The following table provides a summary of the discharges for each MDC class and for each year.

```{r, echo=FALSE, include=TRUE}
#overall discharges for acuti - regime diurno for the 4 years
overall_diurno <- matching_tables_2.2.11 %>%
  dplyr::select(CLASSE_MDC, DIMISSIONI, ANNO)

overall_diurno %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 2.2')
```

The bar chart below shows the distribution of discharges in *Acuti - Regime diurno* per age category, for each year of analysis.

```{r, echo=FALSE, include=TRUE}

sdo_age_gender2 <- sdo_age_gender %>%
  dplyr::filter(`ATTIVITÀ` == "Acuti - Regime diurno") %>%
  dplyr::mutate_at(vars(`Categoria età`), factor)

sdo_age_gender2$`Categoria età` <- factor(sdo_age_gender2$`Categoria età`,
                                          c("Meno di 1 anno",
                                            "Da 1 a 4 anni",
                                            "Da 5 a 9 anni",
                                            "Da 10 a 14 anni",
                                            "Da 15 a 19 anni",
                                            "Da 20 a 24 anni",
                                            "Da 25 a 29 anni", 
                                            "Da 30 a 34 anni", 
                                            "Da 35 a 39 anni",
                                            "Da 40 a 44 anni",
                                            "Da 45 a 49 anni", 
                                            "Da 50 a 54 anni", 
                                            "Da 55 a 59 anni", 
                                            "Da 60 a 64 anni", 
                                            "Da 65 a 69 anni", 
                                            "Da 70 a 74 anni",
                                            "75 anni e oltre"))
sdo_age_gender2 %>%
  dplyr::select(`Categoria età`, MASCHI, FEMMINE, ANNO) %>%
  dplyr::mutate(DIMISSIONI_TOT = MASCHI + FEMMINE) %>%
  dplyr::select(`Categoria età`, DIMISSIONI_TOT, ANNO) %>%
 
  ggplot(., aes(x = `Categoria età`,y = DIMISSIONI_TOT, fill = ANNO)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5))

  
  




```

### ACCESSI {.tabset}

The following graph shows the [*Accessi*](#glossario) for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_diurno %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(Accessi = sum(Accessi)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = Accessi,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "Accessi",
              subtitle = "Acuti - Regime diurno")


```



#### 2019 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       x = "CLASSE MDC",
       y = "Accessi") 
  

```


#### 2018 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       x = "CLASSE MDC",
       y = "Accessi") 
  

```



#### 2017 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       x = "CLASSE MDC",
       y = "Accessi") 
  

```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}
sum_accessi <- SDO_diurno %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3))

ggplot(data = sum_accessi, mapping = aes(x =    forcats::fct_reorder(CLASSE_MDC,                                                               MEAN_accessi), 
      y = MEAN_accessi)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER OGNI MDC",
              subtitle = "Acuti - Regime diurno",
       x = "CLASSE MDC",
       y = "Accessi") 
  


```




### {-}

Table 2.3 provides a summary of Max, Min and Mean values for *"Accessi"* for each MDC.
```{r, echo=FALSE, include=TRUE}

SDO_diurno %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>% 
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
  dplyr::summarise(MAX_accessi = max(`N. max. Accessi.`),
            MIN_accessi = min(`N. min. Accessi`),
            MEAN_accessi = round(mean(`N. medio Accessi`), digits = 3)) %>% 
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 2.3'
    
  )
```


## Riabilitazione - Regime ordinario.

### DIMISSIONI

Table 3.1 provides a summary of the hospedalization discharges in *"Riabilitazione - Regime ordinario"* and the variation from year to year.

```{r, echo=FALSE, include=TRUE}
summary_ordinario_riab <- SDO_ordinario_riab %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_ordinario_riab %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 3.1')



```

This bar chart shows the hospital discharges by MDC for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "HOSPITAL DISCHARGES BY MDC",
              subtitle = "Riabilitazione - Regime ordinario")

  
```

There are just a few MDC classes with a considerable number of discharges in *Riabilitazione - Regime ordinario*. These are: **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**; **MDC 01 - Malattie e disturbi del sistema nervoso**; **MDC 05 - Malattie e disturbi dell'apparato cardiocircolatorio**.

### GIORNATE DI DEGENZA {.tabset}

The following graph shows the *Giorni di Degenza* for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_ordinario_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(GIORNI_DEGENZA = sum(`GIORNATE DI DEGENZA`)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = GIORNI_DEGENZA,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "GIORNI DI DEGENZA",
              subtitle = "Riabilitazione - Regime ordinario")


```


Table 3.2 provides a summary for *Giorni di Degenza* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_ordinario_degenza_riab <- SDO_ordinario_riab %>%
  dplyr::select(CLASSE_MDC, `GIORNATE DI DEGENZA`, ANNO)

overall_ordinario_degenza_riab %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 3.3"
                )

```



The following graph shows the Mean value for *Giorni di Degenza* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2017 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Riabilitazione - Regime ordinario",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

### {-}

Table 3.3 provides a summary for the mean value for *"Giorni di Degenza"* for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_ordinario_riab %>%
  dplyr::select(CLASSE_MDC, `DEGENZA MEDIA (giorni)`, ANNO) %>%
  mutate_at(vars(`DEGENZA MEDIA (giorni)`), funs(round), digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 3.4'
    
  )
  


```


## Riabilitazione - Regime diurno.

### DIMISSIONI

Table 4.1 provides a summary of the hospedalization discharges in *"Riabilitazione - Regime diurno"* and the variation from year to year.

```{r, echo=FALSE, include=TRUE}
summary_diurno_riab <- SDO_diurno_riab %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_diurno_riab %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 4.1')


```



This bar chart shows the hospital discharges by MDC for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "HOSPITAL DISCHARGES BY MDC",
              subtitle = "Riabilitazione - Regime diurno")

  
```

In the case of activity *Riabilitazione - Regime diurno*, there are a few MDC with a considerable level of discharges: **MDC 01 - Malattie e disturbi del sistema nervoso**;
**MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**; **MDC 23 - 23 - Fattori che influenzano lo stato di salute ed il ricorso ai servizi sanitari**.

### ACCESSI {.tabset}

The following graph shows the *Accessi* for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_diurno_riab %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(Accessi = sum(ACCESSI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = Accessi,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "Accessi",
              subtitle = "Riabilitazione - Regime diurno")


```


Table 4.2 provides a summary for *Accessi* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_diurno_accessi_riab <- SDO_diurno_riab %>%
  dplyr::select(CLASSE_MDC, ACCESSI, ANNO)

overall_diurno_accessi_riab %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 4.2"
                )

```



The following graph shows the Mean value for *Accessi* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```


#### 2017 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```



#### 2016 {-}
```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `NUMERO MEDIO DI ACCESSI`), 
                    y = `NUMERO MEDIO DI ACCESSI`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "ACCESSI MEDI PER MDC",
              subtitle = "Riabilitazione - Regime diurno",
       x = "CLASSE MDC",
       y = "ACCESSI") 
  




```


### {-}


Table 4.3 provides a summary for the mean value for *Accessi* for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_diurno_riab %>%
  dplyr::select(CLASSE_MDC, `NUMERO MEDIO DI ACCESSI`, ANNO) %>%
  mutate_at(vars(`NUMERO MEDIO DI ACCESSI`), funs(round), digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 4.3'
    
  )
  


```


## Lungodegenza.

### DIMISSIONI

Table 5.1 provides a summary of the hospedalization discharges in [*Lungodegenza*](#glossario) and the variation from year to year.

```{r, echo=FALSE, include=TRUE}
summary_lungodeg <- SDO_lungodegenza %>% 
  dplyr::group_by(ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate("VARIAZIONE %" = round((DIMISSIONI / lag(DIMISSIONI) -1) * 100, digits = 3))

summary_ordinario_riab %>% 
  DT::datatable( options = list(
    lengthChange = FALSE,
    paging = FALSE,
    searching = FALSE
    
  ),
    caption = 'Table 5.1')



```

This bar chart shows the hospital discharges by MDC for each year of the analysis.
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(DIMISSIONI = sum(DIMISSIONI)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = DIMISSIONI,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "HOSPITAL DISCHARGES BY MDC",
              subtitle = "Lungodegenza")

  
```

As shown from the bar chart, over the 4 years of analysis, the MDC class with the greatest number of dicharges for the activity *Lungodegenza* are: **MDC 08 - Malattie e disturbi del sistema muscolo-scheletrico e del tessuto connettivo**; **MDC 01 - Malattie e disturbi del sistema nervoso**; **MDC 04 - 04 - Malattie e disturbi dell'apparato respiratorio**.

### GIORNATE DI DEGENZA {.tabset}

The following graph shows the *Giorni di Degenza* for each MDC and over the years of analysis.
```{r, echo=FALSE, include=TRUE}
SDO_lungodegenza %>%
  dplyr::group_by(CLASSE_MDC, ANNO) %>%
  dplyr::summarise(GIORNI_DEGENZA = sum(`GIORNATE DI DEGENZA`)) %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  ggplot(.) +
  geom_bar(
    mapping = aes(y = GIORNI_DEGENZA,
                  x = CLASSE_MDC,
                  fill = ANNO),
    position = "dodge",
    width = .5,
           stat = "identity", 
           color = "black") + 
  theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) + 
  labs(title = "GIORNI DI DEGENZA",
              subtitle = "Lungodegenza")


```


Table 5.2 provides a summary for *Giorni di Degenza* by MDC for each year.

```{r, echo=FALSE, include=TRUE}
overall_lungod <- SDO_lungodegenza %>%
  dplyr::select(CLASSE_MDC, `GIORNATE DI DEGENZA`, ANNO)

overall_lungod%>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
                caption = "Table 5.2"
                )

```



The following graph shows the Mean value for *Giorni di Degenza* for each MDC.

#### 2019 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2019") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2018 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2018") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

#### 2017 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2017") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```


#### 2016 {-}
```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::filter(ANNO == "2016") %>%
  dplyr::mutate(CLASSE_MDC = gsub("-.*", " ", CLASSE_MDC)) %>%
  dplyr::mutate(CLASSE_MDC = ifelse(stringr::str_detect(string = CLASSE_MDC,
                                                        pattern = "0|1|2"), 
                                    paste0("MDC ", CLASSE_MDC), CLASSE_MDC)) %>%
  dplyr::mutate_at(vars(ANNO, CLASSE_MDC), factor) %>%
ggplot(mapping = aes(x = forcats::fct_reorder(CLASSE_MDC, 
                                              `DEGENZA MEDIA (giorni)`), 
                    y = `DEGENZA MEDIA (giorni)`)) +
 geom_histogram(stat = "identity") + 
  coord_flip() + 
   theme(axis.text.x  = element_text(angle = 90, 
                                    hjust = 1, 
                                    vjust = 0.5)) +  
  labs(title = "DEGENZA MEDIA PER OGNI MDC",
              subtitle = "Lungodegenza",
       x = "CLASSE MDC",
       y = "GIORNI DI DEGENZA") 
  




```

### {-}

Table 5.3 provides a summary for the mean value for *"Giorni di Degenza"* for each MDC.

```{r, echo=FALSE, include=TRUE}

SDO_lungodegenza %>%
  dplyr::select(CLASSE_MDC, `DEGENZA MEDIA (giorni)`, ANNO) %>%
  mutate_at(vars(`DEGENZA MEDIA (giorni)`), funs(round), digits = 3) %>%
  DT::datatable(
    extensions = 'Buttons', 
                options = list(dom = 'Bfrtip',
                               buttons = c('excel', "csv")
                               ),
    caption = 'Table 5.3'
    
  )
  


```


\pagebreak

# GLOSSARIO {-}

**MDC**: MAJOR DIAGNOSTIC CATEGORY (MDC)
Categoria diagnostica principale: il sistema DRG prevede un livello superiore di raggruppamento basato su un criterio di rilevanza clinica (anatomico o eziologico): in base alla diagnosi principale riportata nella Scheda di Dimissione Ospedaliera viene assegnata la categoria MDC; successivamente, viene attribuito uno specifico DRG fra quelli appartenenti alla MDC selezionata. 

**Giorno di Degenza**: Periodo di 24 ore durante il quale un posto letto è occupato.  

**Accessi**: Presenza in ospedale senza pernottamento, per l’erogazione di prestazioni in regime di day hospital.

**Ricoveri di Lungodegenza**: Sono i ricoveri di pazienti dimessi da reparti appartenenti alla disciplina LUNGODEGENTI (codice 60 nei
modelli di rilevazione del Sistema informativo sanitario). 
